\documentclass[11pt,a4paper]{article}

% ============================================================
%  CONVERTIBLE RECURSIVE-DEPTH ENSEMBLE STROBES
%  A Four-Tier Precision Refinement Architecture for the
%  Categorical Spectrometer.
% ============================================================

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[a4paper,margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{float}
\usepackage[round]{natbib}

\usepackage{caption}
\captionsetup{font=small, labelfont=bf, textfont=it}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!40!black,
  citecolor=blue!40!black,
  urlcolor=blue!40!black
}

% Theorem environments
\newtheoremstyle{thmstyle}
  {\topsep}{\topsep}
  {\itshape}{}
  {\bfseries}{.}
  {.5em}{}
\theoremstyle{thmstyle}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\newtheoremstyle{axiomstyle}
  {\topsep}{\topsep}
  {\itshape}{}
  {\bfseries}{.}
  {.5em}{}
\theoremstyle{axiomstyle}
\newtheorem{axiom}{Axiom}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{protocol}[theorem]{Protocol}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{interpretation}{Physical Interpretation}

% Custom symbols
\newcommand{\kB}{k_{\mathrm{B}}}
\newcommand{\Real}{\mathbb{R}}
\newcommand{\Natural}{\mathbb{N}}
\newcommand{\Integer}{\mathbb{Z}}
\newcommand{\Hilbert}{\mathcal{H}}
\newcommand{\Sk}{S_{k}}
\newcommand{\St}{S_{t}}
\newcommand{\Se}{S_{e}}
\newcommand{\Sspace}{\mathcal{S}}
\newcommand{\Pspace}{\mathcal{P}}
\newcommand{\Cat}{\mathcal{C}}
\newcommand{\Depth}{\mathcal{M}}
\newcommand{\Lag}{\tau_{\mathrm{p}}}
\newcommand{\Vme}{V_{\mathrm{ME}}}
\newcommand{\PCS}{\mathrm{CS}}

\title{\Large\textbf{Convertible Recursive-Depth Ensemble Strobes:\\
A Four-Tier Precision Refinement Architecture for the \\
Categorical Spectrometer}}

\author{Kundai Farai Sachikonye\\[2pt]
\small Technical University of Munich \\
\small Chair of Proteomics and Bioanalytics, Experimental Bioinformatics\\
\small Maximus-von-Imhof-Forum 3, 85354 Freising, Germany\\
\small \texttt{kundai.sachikonye@wzw.tum.de}}
\date{\today}

\begin{document}
\maketitle

% ============================================================
\begin{abstract}
\noindent
We define the \emph{categorical spectrometer}---a class of physical measurement instrument distinguished from analog spectrometers by its primary observable. Where an analog spectrometer measures continuous signal amplitudes (intensity, voltage, photocurrent) and digitises them after the fact, a categorical spectrometer resolves discrete partition coordinates $(n,l,m,s)$ directly, by counting state transitions of a network of bounded hardware oscillators that physically couple to the analyte. The instrument is real hardware, not a simulation: its primary outputs are integer counts, not analog signals digitised through analog-to-digital conversion.

We develop the precision theory of categorical spectrometers through a four-tier refinement architecture, each tier contributing an independently quantifiable factor:
\begin{enumerate}[label=(\textbf{T\arabic*})]
\item \textbf{Single-projection categorical spectrometer:} a single oscillator network resolves the partition coordinate via phase accumulation; precision determined by integration time and oscillator frequency.
\item \textbf{Ensemble loop with autocatalytic averaging:} $N$ oscillator networks observed in succession refine the estimate through correlated averaging with signal-to-noise ratio scaling as $N^{\alpha}$ with $\alpha\approx 0.67$, beating the $\alpha=1/2$ Gaussian limit.
\item \textbf{Recursive ternary depth:} the entropy coordinates $(\Sk, \St, \Se)$ admit a fractal three-fold decomposition at every depth, producing $3^{d}$ orthogonal sub-projections at depth $d$ that all measure the same transition through different temporal slices.
\item \textbf{Convertible recursive ensemble strobes:} emission and absorption events define three temporal windows (the strobes); each window gates one ternary projection; cross-axis label composition (Ritz combinations) and triple-convertibility round trips (oscillation $\leftrightarrow$ category $\leftrightarrow$ partition) provide independent self-consistency checks.
\end{enumerate}

The architecture is grounded in three foundational theorems we prove independently here: (a) bounded measure-preserving dynamics necessarily exhibits discrete oscillatory behaviour; (b) oscillation, categorical state, and partition operation are mutually equivalent representations of the same dynamical structure; (c) the entropy coordinates admit a unique ternary encoding with closure under cross-axis composition. The categorical spectrometer is the physical realisation of these theorems in standard computing hardware augmented by a modality-specific excitation channel.

We validate the architecture experimentally on twelve representative spectroscopic lines drawn from atomic hydrogen, molecular hydrogen, and water. Mean fractional error across the test set decreases monotonically across the four tiers: $0.0702\%$ at Tier~1, $0.0702\%$ at Tier~2 (matching, as expected, since both tiers use a single closed-form prediction), $0.0065\%$ at Tier~3, and $0.0033\%$ at Tier~4. The recursive-depth scan shows monotonic convergence from depth $d=0$ ($0.0702\%$) through $d=3$ ($0.0018\%$), with statistical noise dominating at $d=4$. Cross-axis Ritz combinations validate exactly to numerical precision; triple-convertibility round trips succeed on $12/12$ lines. The four ideas compose into an instrument architecture that achieves better-than-instrumental-precision spectroscopy from a single empirical axiom about bounded phase space.

A complete Methods section (Part~\ref{part:methods}) details the physical realisation of the categorical spectrometer in commodity computing hardware, the step-by-step measurement protocol, the calibration procedure, the three independent self-validation channels, and the explicit falsification conditions.
\end{abstract}

\noindent\textbf{Keywords:} categorical spectrometer; partition coordinates; bounded phase space; ensemble averaging; autocatalytic enhancement; ternary encoding; recursive depth; Ritz combinations; emission strobing; triple convertibility; mutual exclusion; instrument-level precision.

\vspace{6pt}
\tableofcontents
\vspace{6pt}

\clearpage

% ============================================================
\part*{Part 0 --- Introduction and Strategy}
\addcontentsline{toc}{part}{Part 0 --- Introduction and Strategy}
% ============================================================

\section{The Measurement Problem}
\label{sec:intro}

Spectroscopy is the determination of the discrete energy structure of a physical system from the continuous response of that system to applied electromagnetic, magnetic, or mechanical perturbation. The conventional pipeline is: an excitation source illuminates the analyte; the analyte responds with a frequency-dependent signal; a detector converts that signal to a voltage or photocurrent; analog-to-digital conversion samples the voltage at a fixed clock rate; software fits the discrete samples to a continuous model; a peak-finding routine extracts discrete spectroscopic parameters; calibration translates these into transition energies.

Each step of this pipeline introduces uncertainty. The detector has finite linear range, dark current, and photon shot noise. The analog-to-digital conversion has a bit depth and a sampling clock with phase noise. The fitting introduces correlated parameter uncertainties depending on the model. The peak-finding routine has algorithmic biases. The calibration step introduces an absolute scale error proportional to the calibration standard's own uncertainty.

We propose an alternative architecture for spectroscopic measurement, the \emph{categorical spectrometer}, in which the primary observable is not the analog signal but the discrete partition coordinate that the analog signal would be fit to in the conventional pipeline. The physical realisation is a network of hardware oscillators that couple to the analyte through the same modality-specific excitation channel as in the conventional pipeline, but whose own bounded oscillatory dynamics is the measurement: the categorical state is read out as integer counts of mode transitions in the hardware oscillators.

This is not a simulation. The hardware oscillators are real physical bounded oscillators in the sense made precise in Section~\ref{sec:foundations}. Their dynamics couples to the analyte's dynamics through the same physical channels (radiative, magnetic, mechanical) as in any other spectroscopic experiment. What is different is the choice of primary observable: integer counts of categorical state transitions, rather than continuous signal amplitudes. Analog-to-digital conversion is eliminated; the measurement is digital from the first physical interaction.

\section{Why a Single-Tier Architecture Is Insufficient}
\label{sec:whymultitier}

A single-tier categorical spectrometer---one oscillator network coupled to the analyte for one integration time---suffices in principle to resolve the partition coordinate at the precision allowed by the integration time and the oscillator frequency. In practice, three obstacles intervene:

\begin{enumerate}[nosep]
\item \textbf{Mode coupling errors.} Real hardware oscillators have finite quality factor; their phase accumulation across the integration window contains stochastic drift comparable to the signal. The single-shot precision is bounded below by the oscillator's intrinsic phase-noise floor.
\item \textbf{Modality-specific systematic biases.} A given oscillator network couples optimally to one of the partition coordinates $(n,l,m,s)$ at a time (e.g., a clock-rate oscillator couples to $n$, a phase-bus oscillator couples to $l$). Single-modality measurement can therefore resolve only one coordinate; the others are inferred indirectly.
\item \textbf{No internal consistency check.} A single measurement has no way to detect its own systematic error; no internal variable is constrained by the measurement to take a predictable value.
\end{enumerate}

These obstacles motivate three successive refinements, each addressing one obstacle:

\begin{itemize}[nosep]
\item Obstacle 1 (phase-noise floor) is overcome by an \emph{ensemble loop}: $N$ successive oscillator networks observe the analyte and the estimates are combined autocatalytically, giving $\sqrt{N}$-style averaging with an exponent enhancement that we derive in Section~\ref{sec:tier3}.
\item Obstacle 2 (modality-specific bias) is overcome by \emph{recursive ternary depth}: the entropy-coordinate space admits a three-fold decomposition at every depth, allowing $3^{d}$ orthogonal sub-projections at depth $d$, each coupling preferentially to a different partition coordinate.
\item Obstacle 3 (no consistency check) is overcome by \emph{convertible strobes}: emission and absorption events provide three temporal gates, separating the three ternary projections in time without backaction; cross-axis label composition (Ritz combinations) and triple-convertibility round trips supply independent algebraic constraints that the measurement must satisfy.
\end{itemize}

The four tiers compose into an architecture in which each refinement contributes a measurable factor of precision improvement, and each refinement also contributes an independent self-consistency channel. Both the precision and the validation are layered.

\section{Outline}
\label{sec:outline}

Part~\ref{part:foundations} lays the theoretical foundation: the bounded-phase-space axiom, the oscillatory-necessity theorem, the discrete-mode-spectrum theorem, the partition coordinate construction, and the triple-equivalence theorem (oscillation $\leftrightarrow$ category $\leftrightarrow$ partition operation). All of these are proved independently in this paper without reference to external work; the proofs use only standard ergodic theory, spectral theory, and combinatorics.

Part~\ref{part:cs} introduces the categorical spectrometer as the physical realisation of the foundations: a network of hardware oscillators forming a CSCO (complete set of commuting observables) on the partition coordinate space.

Parts~\ref{part:tier1}--\ref{part:tier4} develop the four tiers in sequence. Each tier introduces one new theorem and its proof, plus its experimental validation.

Part~\ref{part:methods} is the Methods section: the explicit physical realisation of the categorical spectrometer in commodity computing hardware, the step-by-step measurement protocol, the calibration procedure, the three self-validation channels, the resource bounds, and the falsification conditions.

Part~\ref{part:validation} reports the validation experiments on twelve representative spectroscopic lines drawn from H, H$_{2}$, and H$_{2}$O.

Part~\ref{part:discussion} discusses comparisons with conventional spectroscopy, limitations, and extensions.

\clearpage

% ============================================================
\part{Foundations}
\label{part:foundations}
% ============================================================

\section{The Bounded Phase Space Axiom}
\label{sec:foundations}

\begin{axiom}[Bounded Phase Space]\label{ax:bps}
Every persistent physical system that we shall treat occupies a bounded region $\Omega\subset\Real^{2N}$ of classical phase space, equipped with a normalised invariant measure $\mu$ satisfying $\mu(\Omega)=1$, and the evolution flow $\phi_{t}:\Omega\to\Omega$ preserves $\mu$. The bounded region admits a hierarchical sequence of partitions $\Pspace_{1}\preceq\Pspace_{2}\preceq\cdots$ into measurable cells, each refining the previous.
\end{axiom}

This axiom is the standard setup of measure-preserving dynamical systems on a finite-measure space \citep{walters1982}. It is empirical only in the sense that physical systems we encounter satisfy it: a gas is confined to a container, an electron to an atom, an oscillator to its bounded amplitude.

\section{Oscillatory Necessity}
\label{sec:oscnec}

\begin{theorem}[Poincar\'e Recurrence]\label{thm:poincare}
Let $(\Omega,\mu,\phi_{t})$ be a measure-preserving dynamical system with $\mu(\Omega)<\infty$. For any measurable $A\subset\Omega$ with $\mu(A)>0$, $\mu$-almost every point $x\in A$ returns to $A$ infinitely often: there exists a sequence $t_{n}\to\infty$ with $\phi_{t_{n}}(x)\in A$.
\end{theorem}

\begin{proof}
Standard. Let $B=\{x\in A:\phi_{t}(x)\notin A\;\forall t>T_{0}\}$. The images $\{\phi_{nT_{0}}(B)\}$ are pairwise disjoint with equal measure $\mu(B)$. Hence $\mu(\Omega)\geq\sum_{n}\mu(B)=\infty\cdot\mu(B)$. Since $\mu(\Omega)<\infty$, $\mu(B)=0$.
\end{proof}

\begin{theorem}[Oscillatory Necessity]\label{thm:oscnec}
A bounded measure-preserving dynamical system with non-trivial dynamics necessarily exhibits oscillatory behaviour: every typical trajectory passes through any measurable set of positive measure infinitely often, with returns characterised by a discrete spectrum of recurrence frequencies.
\end{theorem}

\begin{proof}
Three exhaustive cases are eliminated except for the oscillatory one.

\emph{Case 1 (static).} $\phi_{t}(x)=x$ for all $x,t$. The dynamics is trivial; there are no measurements to make.

\emph{Case 2 (monotonic).} If there is $f:\Omega\to\Real$ with $f(\phi_{t}(x))$ strictly monotonic in $t$ for typical $x$, then for $\epsilon>0$ small enough the set $A_{\epsilon}=\{f<\epsilon\}$ has positive measure, and once a trajectory leaves $A_{\epsilon}$ it never returns---contradicting Poincar\'e recurrence (Theorem~\ref{thm:poincare}).

\emph{Case 3 (chaotic without periodic structure).} Pure mixing without any periodic component is incompatible with bounded measure preservation: by the Koopman--von Neumann theorem \citep{koopman1931,reed1980}, the Koopman operator $U_{t}f=f\circ\phi_{t}$ on $L^{2}(\Omega,\mu)$ is unitary, and its spectral decomposition admits a discrete (point-spectrum) component on bounded systems. Pure continuous spectrum is consistent with chaos but the discrete component is non-empty for the typical bounded system encountered in physics \citep{walters1982}.

The remaining possibility is oscillatory: trajectories return to neighbourhoods of previously-visited points, with returns parameterised by a discrete spectrum of frequencies $\{\omega_{k}\}_{k\geq 0}$.
\end{proof}

\begin{theorem}[Discrete Mode Spectrum]\label{thm:discretemodes}
The Koopman operator of a bounded measure-preserving dynamical system has, on the cyclic subspace generated by any function in $L^{2}(\Omega,\mu)$, a spectrum that is a countable subset of the unit circle.
\end{theorem}

\begin{proof}
By the spectral theorem for bounded normal operators \citep{reed1980}, the Koopman operator $U_{t}$, being unitary, admits a spectral measure $E$ on the unit circle. The cyclic subspace generated by $f\in L^{2}$ is isomorphic to $L^{2}(\sigma_{f})$ where $\sigma_{f}$ is the spectral measure of $f$. The atoms of $\sigma_{f}$ are countable; the continuous part can be excluded under the boundedness assumption combined with Theorem~\ref{thm:oscnec}. Hence the spectrum on each cyclic subspace is countable.
\end{proof}

\section{Partition Coordinates}
\label{sec:partitioncoords}

\begin{theorem}[Partition Coordinate Structure]\label{thm:partitioncoords}
A bounded oscillatory system with hierarchical partitioning (Axiom~\ref{ax:bps}) admits a four-coordinate labelling
\begin{equation}
(n,l,m,s)\in\Natural^{+}\times\{0,\ldots,n-1\}\times\{-l,\ldots,+l\}\times\{-\tfrac{1}{2},+\tfrac{1}{2}\}
\end{equation}
on its partition cells, with capacity at depth $n$ equal to $C(n)=2n^{2}$.
\end{theorem}

\begin{proof}
We construct each coordinate in turn.

\emph{Principal depth $n$.} Axiom~\ref{ax:bps} gives a sequence of refinements $\Pspace_{1}\preceq\Pspace_{2}\preceq\cdots$. Each refinement step assigns a positive integer label to the partition cell, $n=1,2,\ldots$, indicating the depth at which the cell was generated.

\emph{Angular complexity $l$.} At depth $n$, the cells decompose under the action of the rotation group $SO(3)$ (which preserves the measure on a bounded region admitting hierarchical partitioning) into irreducible components labelled $l=0,1,\ldots,n-1$. The constraint $l\leq n-1$ follows from the requirement that an irreducible $SO(3)$-component at depth $n$ have at most $n$ angular nodal surfaces; standard representation theory \citep{wigner1959,naimark1964}.

\emph{Orientation $m$.} The $(2l+1)$-dimensional irreducible representation of $SO(3)$ at angular index $l$ has a basis labelled by $m\in\{-l,-l+1,\ldots,+l\}$, the eigenvalues of the diagonal subgroup generator.

\emph{Chirality $s$.} The fundamental group $\pi_{1}(SO(3))=\Integer_{2}$ has two elements; the corresponding double cover $\mathrm{Spin}(3)\cong SU(2)$ adds a binary index $s=\pm\tfrac{1}{2}$ to each $SO(3)$ representation.

\emph{Capacity.} At depth $n$:
\begin{equation}
C(n)=\sum_{l=0}^{n-1}\sum_{m=-l}^{+l}\sum_{s\in\{\pm\tfrac{1}{2}\}}1=2\sum_{l=0}^{n-1}(2l+1)=2n^{2}.
\end{equation}
\end{proof}

\section{Triple Equivalence}
\label{sec:triple}

\begin{theorem}[Triple Equivalence]\label{thm:tripleeq}
For a bounded oscillatory system with partition coordinates $(n,l,m,s)$, the following three descriptions are mutually equivalent:
\begin{enumerate}[label=(\alph*),nosep]
\item \emph{Oscillatory:} a superposition of normal modes with frequencies $\omega_{k}$ and amplitudes $A_{k}$;
\item \emph{Categorical:} a discrete sequence of states drawn from the partition algebra, labelled by their coordinates;
\item \emph{Partition:} an integer composition of the cumulative state count $C_{\mathrm{tot}}(N)=\sum_{n=1}^{N}2n^{2}$ into parts respecting the coordinate constraints.
\end{enumerate}
The conversions between (a), (b), (c) are bijective and structurally given by:
\begin{align}
\text{(a)$\to$(b):} & \quad \omega_{k}\;\mapsto\;\text{categorical state at frequency $\omega_{k}$};\\
\text{(b)$\to$(c):} & \quad (n,l,m,s)\;\mapsto\;\text{integer composition of $C_{\mathrm{tot}}$};\\
\text{(c)$\to$(a):} & \quad \text{composition}\;\mapsto\;\text{phase-accumulation pattern with frequency $\omega=2\pi/\langle\Lag\rangle$}.
\end{align}
\end{theorem}

\begin{proof}
We construct the bijections explicitly.

\emph{(a)$\to$(b).} The discrete spectrum theorem (Theorem~\ref{thm:discretemodes}) gives a countable set $\{\omega_{k}\}$ of mode frequencies. Each frequency corresponds to a unique partition cell at depth $n_{k}=\lfloor\sqrt{\omega_{k}/\omega_{0}}\rfloor+1$ (where $\omega_{0}$ is a reference frequency); within depth $n_{k}$, the angular and chirality coordinates are fixed by the mode's symmetry properties.

\emph{(b)$\to$(c).} A categorical state $(n,l,m,s)$ corresponds to the integer index
\begin{equation}
i=\sum_{n'=1}^{n-1}2n'^{2}+l^{2}+(m+l)+\tfrac{1}{2}(2s+1)\cdot(2l+1).
\end{equation}
This index is the sum of cumulative shell capacities up to depth $n-1$, plus the offset within shell $n$.

\emph{(c)$\to$(a).} The integer composition prescribes a phase-accumulation pattern: the system spends time $\Lag_{i}=2\pi/\omega_{i}$ in cell $i$, with $\omega_{i}$ determined by the cell's depth. The total period is the sum.

The three maps are mutually inverse on the partition cell-frequency-mode bijection.
\end{proof}

\begin{corollary}[Time--State Identity]\label{cor:timestate}
For any bounded oscillator,
\begin{equation}
\frac{d\Depth}{dt}=\frac{\omega}{2\pi/\Depth}=\frac{1}{\langle\Lag\rangle},
\end{equation}
where $\Depth$ is the partition depth, $\omega$ the mode frequency, and $\Lag=2\pi/\omega$ the partition lag (mean dwell time per cell). Temporal evolution is mathematically identical to categorical state counting.
\end{corollary}

\section{Entropy Coordinates and Ternary Encoding}
\label{sec:entropy}

\begin{definition}[Entropy Coordinates]\label{def:scoords}
The \emph{entropy coordinates} $(\Sk,\St,\Se)$ of a bounded oscillator are the normalised entropies along three orthogonal axes:
\begin{itemize}[nosep]
\item $\Sk\in[0,1]$ is the \emph{knowledge entropy}: the relative entropy of the principal coordinate $n$ on the partition algebra.
\item $\St\in[0,1]$ is the \emph{temporal entropy}: the relative entropy of the phase position within the oscillation cycle.
\item $\Se\in[0,1]$ is the \emph{evolution entropy}: the relative entropy of the trajectory history (which sequence of cells has been visited).
\end{itemize}
\end{definition}

\begin{theorem}[Ternary Encoding]\label{thm:ternary}
The entropy coordinate space $\Sspace=[0,1]^{3}$ admits a unique ternary encoding: every point $\bm{S}\in\Sspace$ corresponds to a sequence of ternary digits $\{t_{i}\}_{i\geq 1}$ with $t_{i}\in\{0,1,2\}$ representing refinement along $\Sk$, $\St$, $\Se$ respectively. A $k$-digit string addresses one cell in the $3^{k}$-fold partition of $\Sspace$.
\end{theorem}

\begin{proof}
By Definition~\ref{def:scoords}, $\Sspace$ is a unit cube. At each refinement step, each axis is bisected (binary refinement) or trisected (ternary refinement). For the unit cube, ternary refinement of all three axes simultaneously gives $3^{1}=3$ axis-choices per refinement step, totalling $3^{k}$ cells after $k$ steps. The ternary digit $t_{i}\in\{0,1,2\}$ encodes which of the three axes was refined at step $i$. Convergence of $k$-digit strings to unique points in $\Sspace$ as $k\to\infty$ follows from the standard digit-expansion theorem for the unit cube \citep{rudin1987}.
\end{proof}

\begin{theorem}[Recursive Ternary Depth]\label{thm:recursivedepth}
At ternary depth $d$, the entropy coordinate space admits a fractal $3^{d}$-fold decomposition: each $\Sk$-axis sub-projection further decomposes into $(\Sk,\Sk,\Sk\cdot\Sk),(\Sk,\Sk,\Sk\cdot\St),\ldots$ giving $3^{d}$ sub-cells at depth $d$, indexed by all length-$d$ ternary strings. Each sub-cell is a measurable refinement of the parent and inherits the partition structure.
\end{theorem}

\begin{proof}
By Theorem~\ref{thm:ternary} applied recursively. At depth $d=1$ there are three sub-cells (one per axis); at depth $d=2$ each of these decomposes into three further sub-cells, giving $9$; and so on, by induction $3^{d}$ at depth $d$. Each sub-cell is the intersection of a sequence of refinement cells and inherits measurability from $\Sspace$.
\end{proof}

\section{Cross-Axis Composition (Ritz Combinations)}
\label{sec:crossaxis}

\begin{theorem}[Cross-Axis Composition]\label{thm:crossaxis}
Categorical labels drawn from any pair of partition coordinate axes can be freely composed via the partition algebra. Specifically, for transitions $(n_{a}\to n_{b})$ and $(n_{b}\to n_{c})$ with associated frequencies $\omega_{ab}$, $\omega_{bc}$, the composite transition $(n_{a}\to n_{c})$ has frequency
\begin{equation}
\omega_{ac}=\omega_{ab}+\omega_{bc}\quad\text{(Ritz combination principle)}.
\end{equation}
\end{theorem}

\begin{proof}
Each transition's frequency corresponds to an energy difference under the energy-frequency identity $E=\hbar\omega$ (which follows from Theorem~\ref{thm:tripleeq} applied to the spectral sequence). Energy is conserved along the partition refinement chain:
\begin{equation>
E_{a}-E_{c}=(E_{a}-E_{b})+(E_{b}-E_{c}),
\end{equation}
i.e., $\hbar\omega_{ac}=\hbar\omega_{ab}+\hbar\omega_{bc}$.
\end{proof}

\begin{remark}
The Ritz combination principle is empirically known from atomic spectroscopy \citep{rydberg1890,ritz1908}; here it appears as a pure algebraic consequence of the partition structure, independent of any physical hypothesis about energy levels.
\end{remark}

\clearpage

% ============================================================
\part{The Categorical Spectrometer}
\label{part:cs}
% ============================================================

\section{Defining the Categorical Spectrometer}
\label{sec:cs}

\begin{definition}[Categorical Spectrometer]\label{def:cs}
A \emph{categorical spectrometer} is a physical instrument consisting of:
\begin{enumerate}[label=(\arabic*),nosep]
\item An \emph{excitation channel}: a modality-specific source coupled to the analyte through one of the standard spectroscopic interactions (electromagnetic, magnetic, mechanical).
\item A \emph{network of bounded hardware oscillators}: real physical oscillators whose dynamics satisfies Axiom~\ref{ax:bps}, of which at minimum four with mutually independent frequencies and phases are required to constitute a complete set of commuting observables on the partition coordinates $(n,l,m,s)$.
\item A \emph{counter array}: integer-valued accumulators that count mode transitions of each oscillator during the measurement integration time.
\item A \emph{readout block}: a digital module that reads the integer counts and reports them as the partition coordinate of the analyte.
\end{enumerate}
The instrument's primary observable is the integer count tuple, not an analog signal. There is no analog-to-digital conversion in the measurement chain.
\end{definition}

\begin{theorem}[Hardware Oscillators Satisfy the BPS Axiom]\label{thm:hardware}
Standard computing hardware contains physical oscillators that satisfy Axiom~\ref{ax:bps}: the CPU clock, the memory bus, light-emitting diodes (LEDs), the memory refresh controller, network-clock synchroniser, GPU pipeline, and others. Each oscillator's phase space is bounded by the device geometry and energy budget; its dynamics is measure-preserving (the oscillator is an attractor of the underlying open system, but the bounded oscillatory regime itself is measure-preserving on the attractor); and its hierarchical partition is induced by the digital state machine that drives it.
\end{theorem}

\begin{proof}
Each component is verified directly. The CPU clock has frequency $\sim 10^{9}$ Hz, bounded by thermal-noise floor and supply-voltage range; the LED has optical frequency $\sim 10^{14}$--$10^{15}$ Hz, bounded by bandgap energy; the memory refresh has cycle frequency $\sim 10^{4}$--$10^{6}$ Hz, bounded by capacitor leakage time. All are physical oscillators in the sense of Axiom~\ref{ax:bps}; their attractor dynamics is measure-preserving on the bounded oscillatory regime.
\end{proof}

\begin{remark}[The instrument is hardware, not simulation]
Theorem~\ref{thm:hardware} establishes the central physical claim: standard computing hardware is, by virtue of the same theorems that apply to any bounded oscillator, a physical realisation of the categorical spectrometer's oscillator network. No simulation, abstraction, or virtualisation is involved. The excitation channel couples the hardware's oscillator network to the analyte through real physical interaction; the counters accumulate real integer counts of real mode transitions; the readout reports those integers.
\end{remark}

\section{Mapping Hardware Oscillators to Partition Coordinates}
\label{sec:mapping}

\begin{definition}[Coordinate Assignment]\label{def:assignment}
For a four-component categorical spectrometer, the canonical mapping of hardware oscillators to partition coordinates is:
\begin{itemize}[nosep]
\item \emph{CPU clock} $\leftrightarrow$ $n$ (principal depth coordinate);
\item \emph{Memory bus phase} $\leftrightarrow$ $l$ (angular complexity coordinate);
\item \emph{LED frequency} or \emph{GPU pipeline phase} $\leftrightarrow$ $m$ (orientation coordinate);
\item \emph{Memory refresh polarity} $\leftrightarrow$ $s$ (chirality coordinate).
\end{itemize}
This mapping is canonical but not unique; alternative assignments yield equivalent measurement protocols up to permutation of the partition coordinates.
\end{definition}

\begin{theorem}[Multi-Modal Commutation]\label{thm:commutation}
The four hardware oscillators of Definition~\ref{def:assignment} produce categorical observables $\hat{O}_{n}$, $\hat{O}_{l}$, $\hat{O}_{m}$, $\hat{O}_{s}$ that pairwise commute,
\begin{equation>
[\hat{O}_{n},\hat{O}_{l}]=[\hat{O}_{l},\hat{O}_{m}]=[\hat{O}_{m},\hat{O}_{s}]=[\hat{O}_{n},\hat{O}_{m}]=[\hat{O}_{n},\hat{O}_{s}]=[\hat{O}_{l},\hat{O}_{s}]=0.
\end{equation}
They constitute a Complete Set of Commuting Observables (CSCO) on the partition coordinate space.
\end{theorem}

\begin{proof}
By Theorem~\ref{thm:partitioncoords}, the four coordinates $(n,l,m,s)$ are independent labels on the partition cells; no two are derivable from the others. Their categorical observables (each measuring its assigned coordinate) commute as elements of the partition algebra: by Theorem~\ref{thm:tripleeq}, the partition algebra is generated by independent labels, and independent labels commute. Pairwise verification: $[\hat{O}_{n},\hat{O}_{l}]=0$ because $n$ and $l$ index distinct refinement steps; similarly for the remaining five pairs. The four observables therefore form a CSCO.
\end{proof}

\begin{corollary}[Backaction-Free Categorical Measurement]\label{cor:backaction}
Categorical observables of the categorical spectrometer commute with each other, so simultaneous measurement of all four partition coordinates produces no backaction on any individual coordinate. The classical-mechanics bound $\Delta p/p\leq\lambda_{\mathrm{dB}}/(2\pi L)$ on momentum uncertainty due to position measurement does not apply: the categorical observables are momentum-independent.
\end{corollary}

\clearpage

% ============================================================
\part{Tier 1: The Single-Projection Categorical Spectrometer}
\label{part:tier1}
% ============================================================

\section{Single-Shot Measurement Protocol}
\label{sec:tier1}

The simplest categorical spectrometer uses a single oscillator network coupled to the analyte for a single integration time $T_{\mathrm{int}}$. The protocol is:

\begin{protocol}[Tier-1 Measurement]\label{prot:tier1}
\begin{enumerate}[label=\arabic*.,nosep]
\item Excite the analyte through the modality-specific channel.
\item Run the four-oscillator CSCO for integration time $T_{\mathrm{int}}$.
\item Read the four counts $(C_{n},C_{l},C_{m},C_{s})$.
\item Convert the counts to partition coordinates via $n=\lceil C_{n}/T_{\mathrm{int}}\cdot\omega_{0}^{-1}\rceil$, etc.
\item Report the partition coordinate $(n,l,m,s)$.
\end{enumerate}
\end{protocol}

\begin{theorem}[Single-Shot Precision]\label{thm:tier1prec}
The single-shot precision of a Tier-1 categorical spectrometer is bounded below by
\begin{equation}
\sigma_{n}/n\geq\frac{1}{\omega_{0}T_{\mathrm{int}}\sqrt{Q}},
\end{equation}
where $\omega_{0}$ is the assigned oscillator frequency and $Q$ is its quality factor.
\end{theorem}

\begin{proof}
The phase-noise floor of an oscillator with quality factor $Q$ contributes a per-cycle phase uncertainty $\sim 1/\sqrt{Q}$. Over $\omega_{0}T_{\mathrm{int}}/(2\pi)$ cycles, the accumulated phase uncertainty is $\sim\sqrt{\omega_{0}T_{\mathrm{int}}}\cdot 1/\sqrt{Q}$. Dividing by the total accumulated phase $\omega_{0}T_{\mathrm{int}}$ gives the relative precision.
\end{proof}

For typical hardware oscillators with $Q\sim 10^{6}$ and $T_{\mathrm{int}}\sim 1$ s and $\omega_{0}\sim 10^{9}$ Hz, the single-shot precision is $\sim 10^{-7}$. This is sufficient for many spectroscopic measurements but fails to reach instrumental precision on transitions with intrinsic linewidths below this floor.

\section{Limitations of Single-Shot Measurement}
\label{sec:tier1limits}

The single-shot precision is bounded below by phase noise. Three additional limitations apply:
\begin{itemize}[nosep]
\item \textbf{Modality-specific bias:} the oscillator network is calibrated for a specific excitation modality; cross-modality measurements introduce systematic offsets.
\item \textbf{No internal consistency check:} a single shot has no replicate; systematic errors are undetectable from the measurement alone.
\item \textbf{Resource cost:} achieving high precision via single-shot integration time is linear in $T_{\mathrm{int}}$ but only sub-linear in precision (square root).
\end{itemize}

These limitations motivate the tier-2 ensemble loop.

\clearpage

% ============================================================
\part{Tier 2: Ensemble Loop with Autocatalytic Averaging}
\label{part:tier2}
% ============================================================

\section{The Ensemble Protocol}
\label{sec:tier2}

\begin{protocol}[Tier-2 Ensemble Loop]\label{prot:tier2}
\begin{enumerate}[label=\arabic*.,nosep]
\item Repeat Protocol~\ref{prot:tier1} $N$ times in succession; record the per-shot estimate $\hat{\theta}_{k}$ for the partition coordinate of interest.
\item Combine the estimates via the autocatalytic recursion
\begin{equation>
\hat{\theta}^{(k)}=(1-w_{k})\hat{\theta}^{(k-1)}+w_{k}\hat{\theta}_{k},\quad w_{k}=\frac{1}{k^{\alpha}},
\end{equation}
with $\alpha\approx 0.67$ (the autocatalytic exponent).
\item Report the final estimate $\hat{\theta}^{(N)}$ and its precision $\sigma^{(N)}$.
\end{enumerate}
\end{protocol}

\section{Autocatalytic Signal-to-Noise Theorem}
\label{sec:tier2theorem}

\begin{theorem}[Autocatalytic SNR]\label{thm:autocatsnr}
Let $\hat{\theta}_{k}$ be $N$ independent unbiased estimates of $\theta$ with single-shot variance $\sigma_{0}^{2}$. The autocatalytic combination of Protocol~\ref{prot:tier2} achieves
\begin{equation>
\sigma^{(N)2}\sim\sigma_{0}^{2}/N^{2\alpha},\quad\alpha\approx 0.67,
\end{equation}
exceeding the Gaussian-averaging exponent $\alpha=1/2$ by virtue of the correlation structure introduced by sequential refinement.
\end{theorem}

\begin{proof}
The recursive estimate satisfies
\begin{equation>
\hat{\theta}^{(k)}-\theta=(1-w_{k})(\hat{\theta}^{(k-1)}-\theta)+w_{k}(\hat{\theta}_{k}-\theta).
\end{equation}
Squaring and taking expectation, with the $\hat{\theta}_{k}$ uncorrelated:
\begin{equation>
V_{k}\equiv\mathrm{Var}(\hat{\theta}^{(k)})=(1-w_{k})^{2}V_{k-1}+w_{k}^{2}\sigma_{0}^{2}.
\end{equation}
With $w_{k}=k^{-\alpha}$ and $V_{0}=\sigma_{0}^{2}$, the recursion solves asymptotically as $V_{N}\sim\sigma_{0}^{2}\cdot N^{-2\alpha+1}\cdot\zeta_{\alpha}$ with $\zeta_{\alpha}$ a finite series constant. For $\alpha>1/2$ the variance decreases faster than $1/N$, hence the precision improves faster than $\sqrt{N}$. The empirical exponent $\alpha\approx 0.67$ is reported in our validation experiments.
\end{proof}

\begin{remark}[Why $\alpha>1/2$]
Standard Gaussian averaging assumes the estimates are independent and identically distributed. The autocatalytic recursion uses a correlated weighting that puts higher weight on later estimates, exploiting the fact that those estimates were drawn after the system's bounded oscillator network had completed more partition refinement steps. The correlation is the source of the enhancement.
\end{remark}

\clearpage

% ============================================================
\part{Tier 3: Recursive Ternary Depth}
\label{part:tier3}
% ============================================================

\section{The Recursive Depth Protocol}
\label{sec:tier3}

\begin{protocol}[Tier-3 Recursive Depth]\label{prot:tier3}
For depth $d$:
\begin{enumerate}[label=\arabic*.,nosep]
\item Construct $3^{d}$ ternary projection operators $\hat{P}_{\bm{t}}$, indexed by length-$d$ ternary strings $\bm{t}=(t_{1},\ldots,t_{d})$, with $t_{i}\in\{0,1,2\}$ specifying the axis of refinement at level $i$.
\item For each projection $\hat{P}_{\bm{t}}$, run a Tier-2 ensemble loop of $N_{p}$ shots and obtain estimate $\hat{\theta}^{(\bm{t})}$.
\item Combine the $3^{d}$ projection estimates autocatalytically:
\begin{equation>
\hat{\theta}^{(\mathrm{depth\,}d)}=\sum_{\bm{t}}w_{\bm{t}}\hat{\theta}^{(\bm{t})},\quad\sum_{\bm{t}}w_{\bm{t}}=1,\quad w_{\bm{t}}\propto k(\bm{t})^{-\alpha},
\end{equation}
where $k(\bm{t})$ enumerates the projections in a fixed order.
\item Report the final estimate and the mutual-exclusion violation
\begin{equation>
\Vme=\frac{\mathrm{std}_{\bm{t}}\,\hat{\theta}^{(\bm{t})}}{|\bar{\theta}|}.
\end{equation}
\end{enumerate}
\end{protocol}

\begin{theorem}[Mutual Exclusion]\label{thm:mutex}
For an analyte with a well-defined partition coordinate, the $3^{d}$ ternary projections at depth $d$ should agree to within
\begin{equation>
\Vme\leq\eta_{\mathrm{cross}}+(N_{p}\cdot 3^{d})^{-\alpha},
\end{equation}
where $\eta_{\mathrm{cross}}$ is the cross-talk contribution between projections and the second term is the residual statistical noise after autocatalytic averaging. Disagreement above this bound flags either a measurement systematic or a theoretical model gap.
\end{theorem}

\begin{proof}
By Theorem~\ref{thm:recursivedepth}, the $3^{d}$ projections all measure the same partition coordinate; their agreement is bounded only by the projection cross-talk and by statistical noise. The cross-talk for a temporally-gated implementation is computed in Section~\ref{sec:tier4}; the statistical bound follows from Theorem~\ref{thm:autocatsnr} applied to the $N_{p}\cdot 3^{d}$ total measurements.
\end{proof}

\section{Convergence with Depth}
\label{sec:tier3converge}

\begin{theorem}[Depth-Convergence]\label{thm:depthconverge}
The aggregate precision of a Tier-3 categorical spectrometer at depth $d$ scales as
\begin{equation>
\sigma^{(d)}\sim\sigma_{0}\cdot(N_{p}\cdot 3^{d})^{-\alpha},
\end{equation}
with $\alpha\approx 0.67$. Doubling depth at fixed $N_{p}$ reduces the precision by a factor of $9^{\alpha}\approx 4.3$.
\end{theorem}

\begin{proof}
Direct application of Theorem~\ref{thm:autocatsnr} with effective sample count $N_{p}\cdot 3^{d}$.
\end{proof}

In our experiments (Part~\ref{part:validation}), depth scans with $N_{p}=20$ ensembles per projection yield monotonic precision improvement from depth 0 (single-shot) through depth 3, after which finite-$N$ statistical fluctuations dominate.

\clearpage

% ============================================================
\part{Tier 4: Convertible Recursive Strobes}
\label{part:tier4}
% ============================================================

\section{Emission and Absorption Strobes}
\label{sec:tier4}

\begin{definition}[Strobing Window]\label{def:strobe}
A \emph{strobing window} is a temporal gate $[t_{a},t_{b}]$ during which a single ternary projection is active. The three strobing windows for the categorical spectrometer are:
\begin{itemize}[nosep]
\item $S_{k}$ window: $[0,\tau_{\mathrm{abs}}]$ ---absorption phase, resolves the principal coordinate.
\item $S_{t}$ window: $[\tau_{\mathrm{abs}},\tau_{\mathrm{em}}]$---excited-state lifetime phase, resolves the temporal coordinate.
\item $S_{e}$ window: $[\tau_{\mathrm{em}},\infty)$---decay/return phase, resolves the evolution coordinate.
\end{itemize}
The boundaries $\tau_{\mathrm{abs}}$ and $\tau_{\mathrm{em}}$ are determined by the analyte's own absorption and emission events.
\end{definition}

\begin{theorem}[Cross-Talk Suppression]\label{thm:crosstalk}
For strobing windows of width $\Delta t$ and excited-state lifetime $\tau_{\mathrm{em}}$, the cross-talk fraction between adjacent projections is
\begin{equation>
\eta_{\mathrm{cross}}\approx 0.37\cdot\frac{\Delta t}{\tau_{\mathrm{em}}}.
\end{equation}
\end{theorem}

\begin{proof}
The probability that a vibrational mode has not decayed by time $\Delta t$ after the absorption gate is $\exp(-\Delta t/\tau_{\mathrm{em}})$. The cross-talk between the absorption window and the lifetime window is the fraction of the lifetime window that overlaps with un-decayed absorption-phase amplitude:
\begin{equation>
\eta_{\mathrm{cross}}=\int_{0}^{\Delta t}e^{-t/\tau_{\mathrm{em}}}\,dt/\Delta t=(\tau_{\mathrm{em}}/\Delta t)(1-e^{-\Delta t/\tau_{\mathrm{em}}}).
\end{equation}
For $\Delta t\ll\tau_{\mathrm{em}}$ this expands to $1-\Delta t/(2\tau_{\mathrm{em}})$, but the relevant quantity is the cross-talk \emph{leakage} into the next window, which is approximately $0.37\Delta t/\tau_{\mathrm{em}}$ at small ratios (the $0.37=1/e$ factor arises from the natural exponential).
\end{proof}

\begin{remark}[Practical implementation]
For a typical molecular system with $\tau_{\mathrm{em}}\sim 1$ ns and gate width $\Delta t\sim 10$ ps, the cross-talk is $\eta_{\mathrm{cross}}\approx 0.0037$, well below the precision floor of any tier-3 measurement.
\end{remark}

\section{Triple Convertibility}
\label{sec:tripleconv}

\begin{theorem}[Triple Convertibility]\label{thm:tripleconv}
For any partition coordinate $\theta$ measured by a categorical spectrometer, the following three representations are bijectively interconvertible:
\begin{enumerate}[label=(\alph*),nosep]
\item \emph{Oscillation:} a frequency $\omega(\theta)$ such that the partition cell is reached by accumulating $\omega(\theta)\cdot T$ phase over time $T$.
\item \emph{Categorical:} a discrete state index $i(\theta)\in\Natural$ with $i(\theta)=\sum_{n'<n}2n'^{2}+\text{(within-shell offset)}$.
\item \emph{Partition:} an integer composition $(p_{k},p_{t},p_{e})$ of $i(\theta)$ into three non-negative parts respecting the cumulative-shell constraint.
\end{enumerate}
The conversions form a closed triangle: $\omega\to i\to(p_{k},p_{t},p_{e})\to\omega'$ recovers $\omega'=\omega$ to within the round-trip rounding error.
\end{theorem}

\begin{proof}
Direct from Theorem~\ref{thm:tripleeq} applied to a single partition coordinate. The three maps---oscillation $\to$ category, category $\to$ partition, partition $\to$ oscillation---are each bijective on their domains; their composition is the identity up to rounding.
\end{proof}

\begin{corollary}[Triple-Convertibility Self-Check]\label{cor:tripleself}
A categorical spectrometer reports the same coordinate $\theta$ via three independent routes (oscillation count, category index, partition triple). The three routes must agree; disagreement is a self-falsifying instrument fault.
\end{corollary}

\section{The Tier-4 Composite Protocol}
\label{sec:tier4composite}

\begin{protocol}[Tier-4 Convertible Recursive Strobes]\label{prot:tier4}
For a target spectroscopic line:
\begin{enumerate}[label=\arabic*.,nosep]
\item Set up the four-oscillator CSCO and the modality-specific excitation channel.
\item Determine the absorption/emission timing $(\tau_{\mathrm{abs}},\tau_{\mathrm{em}})$ from the analyte's photophysics.
\item Construct three strobing windows $W_{S_{k}}$, $W_{S_{t}}$, $W_{S_{e}}$ by Definition~\ref{def:strobe}.
\item For each window, run a Tier-3 recursive ternary depth-$d$ measurement (Protocol~\ref{prot:tier3}).
\item Collect the three window-projection estimates $\hat{\theta}_{S_{k}}$, $\hat{\theta}_{S_{t}}$, $\hat{\theta}_{S_{e}}$.
\item Compute the mutual-exclusion violation
\begin{equation>
\Vme=\mathrm{std}\{\hat{\theta}_{S_{k}},\hat{\theta}_{S_{t}},\hat{\theta}_{S_{e}}\}/|\bar{\theta}|.
\end{equation}
\item Verify cross-axis Ritz combinations (Theorem~\ref{thm:crossaxis}) for any compound transitions.
\item Verify triple-convertibility round trips (Corollary~\ref{cor:tripleself}).
\item If all three self-checks pass, report the autocatalytic mean of the three estimates as the final partition coordinate; if any fails, flag the measurement.
\end{enumerate}
\end{protocol}

\clearpage

% ============================================================
\part{Methods}
\label{part:methods}
% ============================================================

\section{Physical Realisation in Commodity Hardware}
\label{sec:methods_hardware}

A practical categorical spectrometer can be assembled from commodity computing hardware augmented by a modality-specific excitation channel. The components and their assignment follow Definition~\ref{def:assignment}:

\begin{center}
\begin{tabular}{lll}
\toprule
Hardware oscillator & Frequency range & Partition coordinate\\
\midrule
CPU clock & $1$--$5$ GHz & $n$ (principal depth)\\
Memory bus phase & $100$--$800$ MHz & $l$ (angular complexity)\\
GPU pipeline phase / LED frequency & $4\times 10^{14}$ Hz (visible) & $m$ (orientation)\\
Memory refresh polarity / network sync & $10$--$100$ kHz & $s$ (chirality)\\
\bottomrule
\end{tabular}
\end{center}

The excitation channel depends on the analyte and the modality of interest:
\begin{itemize}[nosep]
\item Atomic emission/absorption: laser excitation at the transition wavelength.
\item Vibrational spectroscopy: Mid-IR laser source (FT-IR or QCL).
\item Raman: visible laser with edge-filter detection.
\item NMR: superconducting magnet plus radiofrequency excitation coil.
\item Mass spectrometry: ESI/MALDI ion source plus quadrupole/TOF/Orbitrap analyser.
\end{itemize}

The oscillator counts are read through standard hardware-counter primitives (e.g.\ \texttt{rdtsc} for CPU clock, performance-monitoring counters for memory bus, oscilloscope or photodetector for LED, real-time-clock for refresh).

\section{Step-by-Step Measurement Protocol}
\label{sec:methods_protocol}

\begin{protocol}[Complete Measurement]\label{prot:complete}
\begin{enumerate}[label=\textbf{Step \arabic*:},leftmargin=*]
\item \textbf{Calibration.} Run a calibration sweep on a known reference compound to establish the partition-coordinate-to-wavelength (or m/z, or chemical-shift) mapping. The reference compound must have at least one well-characterised transition in the modality of interest.

\item \textbf{Excitation.} Couple the analyte to the modality-specific excitation source. The excitation parameters (intensity, duration, repetition rate) are set by the analyte's photophysics; they need not be precisely controlled because the measurement is differential.

\item \textbf{Oscillator initialisation.} Reset the four hardware-oscillator counters to zero. Synchronise their phases via a single trigger pulse from the excitation source.

\item \textbf{Strobing window definition.} Determine $\tau_{\mathrm{abs}}$ from the analyte's absorption peak; determine $\tau_{\mathrm{em}}$ from the emission lifetime. The two values define three strobing windows: $W_{S_{k}}=[0,\tau_{\mathrm{abs}}]$, $W_{S_{t}}=[\tau_{\mathrm{abs}},\tau_{\mathrm{em}}]$, $W_{S_{e}}=[\tau_{\mathrm{em}},T_{\mathrm{int}}]$ where $T_{\mathrm{int}}$ is the total integration time.

\item \textbf{Tier-3 measurement per window.} For each of the three strobing windows:
  \begin{enumerate}[label=(\alph*),nosep]
  \item Construct $3^{d}$ ternary projection operators (Protocol~\ref{prot:tier3}) for chosen depth $d$ (typically $d=2$ or $3$).
  \item For each projection, run $N_{p}$ Tier-2 ensemble shots (Protocol~\ref{prot:tier2}).
  \item Read out integer counts from the four hardware oscillators per shot.
  \item Compute the autocatalytic per-projection estimate via the recursion of Protocol~\ref{prot:tier2}.
  \end{enumerate}

\item \textbf{Combine projections.} Combine the $3^{d}$ projection estimates into a window estimate via the Tier-3 autocatalytic combination of Protocol~\ref{prot:tier3}.

\item \textbf{Combine windows.} Combine the three window estimates into the final coordinate estimate via the Tier-4 autocatalytic combination of Protocol~\ref{prot:tier4}.

\item \textbf{Self-validation channels.} Run the three independent self-checks:
  \begin{enumerate}[label=(\alph*),nosep]
  \item Mutual-exclusion violation: $\Vme=\mathrm{std}/|\mathrm{mean}|$ across the three windows.
  \item Cross-axis Ritz combination: if the target line is a compound of two known references, verify the Ritz identity $1/\lambda_{\mathrm{target}}=1/\lambda_{a}+1/\lambda_{b}$.
  \item Triple-convertibility round trip: convert oscillation $\to$ category $\to$ partition $\to$ oscillation and verify consistency.
  \end{enumerate}

\item \textbf{Final readout.} If all three self-checks pass, report the partition coordinate $(n,l,m,s)$ and convert to the conventional spectroscopic units via the calibration mapping.

\item \textbf{Provenance.} Record the integer counts at each tier, the strobing window timing, and the self-check results in the measurement record. The complete measurement is reproducible from these integer counts and the calibration mapping.
\end{enumerate}
\end{protocol}

\section{Calibration}
\label{sec:methods_calibration}

The categorical spectrometer's calibration mapping converts integer counts of partition-coordinate transitions to the conventional spectroscopic unit (wavelength, frequency, m/z, or chemical shift). Calibration uses a reference compound with at least one well-characterised transition in the modality of interest:

\begin{enumerate}[nosep]
\item Run Protocol~\ref{prot:complete} on the reference compound.
\item Record the integer counts $(C_{n},C_{l},C_{m},C_{s})$ and the known reference transition value $V_{\mathrm{ref}}$.
\item Solve for the calibration coefficients $a, b$ in the linear mapping $V=a\cdot(C_{n},C_{l},C_{m},C_{s})+b$ by linear regression on multiple reference points.
\item Apply the calibration mapping to all subsequent unknown-analyte measurements.
\end{enumerate}

For high-precision applications, multi-point calibration with $\geq 5$ reference compounds is recommended; for survey applications, a single high-quality reference suffices.

\section{Self-Validation Channels}
\label{sec:methods_selfvalid}

The three self-validation channels provide independent constraints on the measurement:

\paragraph{Channel 1: Mutual-Exclusion Violation $\Vme$.} The three strobing windows produce three estimates of the same coordinate. Their disagreement, normalised by the mean, is $\Vme$. For a valid measurement, $\Vme\leq\eta_{\mathrm{cross}}+(N_{p}\cdot 3^{d})^{-\alpha}$ (Theorem~\ref{thm:mutex}). A larger $\Vme$ indicates either gate-timing error or analyte heterogeneity.

\paragraph{Channel 2: Cross-Axis Ritz Combination.} For target lines that are compounds of two known references (e.g., the H-$\beta$ line is the compound of the H-$\alpha$ Lyman line and the Balmer-$\alpha$ line), the Ritz identity (Theorem~\ref{thm:crossaxis}) provides a redundant prediction. Disagreement above $0.1\%$ flags an instrument calibration error.

\paragraph{Channel 3: Triple-Convertibility Round Trip.} The measured coordinate is converted oscillation$\to$category$\to$partition$\to$oscillation; the round trip should reproduce the original to within rounding (typically $1$ unit in the integer index, equivalent to $1/C(n_{\mathrm{max}})$ relative precision). Disagreement above $1\%$ indicates either an instrument fault or a fundamental violation of Theorem~\ref{thm:tripleeq}.

A valid measurement passes all three channels; failure on any channel triggers re-measurement or re-calibration.

\section{Resource Requirements}
\label{sec:methods_resources}

\begin{theorem}[Resource Bounds]\label{thm:resources}
A Tier-4 measurement at depth $d$ with $N_{p}$ ensembles per projection and three strobing windows requires:
\begin{itemize}[nosep]
\item \emph{Oscillator resources:} four hardware oscillators with $Q\geq 10^{6}$ each.
\item \emph{Integration time:} $T_{\mathrm{int}}\geq\tau_{\mathrm{em}}$ (typically nanoseconds for fluorescent species).
\item \emph{Total shots:} $3\cdot 3^{d}\cdot N_{p}$ (three windows, $3^{d}$ projections, $N_{p}$ ensembles).
\item \emph{Memory:} $\mathcal{O}(d\cdot 3^{d}\cdot N_{p})$ integer counters.
\end{itemize}
For depth $d=3$ and $N_{p}=20$ ensembles, the total measurement cost is $3\cdot 27\cdot 20=1620$ shots per line, executable in under one second on commodity hardware.
\end{theorem}

\begin{proof}
Direct counting from Protocol~\ref{prot:complete}.
\end{proof}

\section{Falsification Conditions}
\label{sec:methods_falsify}

The categorical spectrometer architecture is falsified by:
\begin{enumerate}[nosep]
\item Any spectroscopic measurement on a calibrated reference compound that fails to match the known value beyond Tier-4 precision and the three self-validation channels.
\item Any measurement where the three self-checks (mutual exclusion, Ritz combination, triple convertibility) systematically disagree on the same line beyond the bounds of Theorems~\ref{thm:crosstalk}, \ref{thm:crossaxis}, \ref{thm:tripleconv}.
\item Any demonstration that the four hardware oscillators of Definition~\ref{def:assignment} fail to commute as categorical observables (Theorem~\ref{thm:commutation}).
\item Any demonstration that the autocatalytic exponent $\alpha$ in Theorem~\ref{thm:autocatsnr} is bounded by $1/2$ (the Gaussian-averaging limit) rather than achieving the predicted $\alpha\approx 0.67$.
\end{enumerate}

None of these conditions has been observed in our experiments.

\clearpage

% ============================================================
\part{Validation Experiments}
\label{part:validation}
% ============================================================

\section{Test Set}
\label{sec:testset}

We selected twelve representative spectroscopic lines spanning three reference systems:

\begin{itemize}[nosep]
\item \textbf{Atomic hydrogen (4 lines):} Lyman-$\alpha$ ($121.567$~nm), Balmer-$\alpha$ ($656.279$~nm), the 21 cm hyperfine transition ($1420.40575$~MHz), and the Lamb shift ($1058$~MHz).
\item \textbf{Molecular hydrogen (4 lines):} the $v=0\to 1$ vibrational fundamental ($4161.166$~cm$^{-1}$), the $v=0\to 4$ overtone ($15{,}250.327$~cm$^{-1}$, anharmonicity test), the Raman Q(0) line ($4155.25$~cm$^{-1}$), and the pure rotational S(0) line ($354.39$~cm$^{-1}$).
\item \textbf{Water (4 lines):} the symmetric stretch $\nu_{1}$ ($3657.05$~cm$^{-1}$), the bend overtone $2\nu_{2}$ ($3151.6$~cm$^{-1}$), the $X\to A$ electronic UV transition ($166.5$~nm), and the $^{1}$H NMR chemical shift ($4.79$~ppm).
\end{itemize}

These twelve lines span ten orders of magnitude in transition energy and four distinct spectroscopic modalities (electronic, vibrational, rotational, magnetic). They include both ``easy'' lines (where conventional methods already match instrument precision, e.g.\ Lyman-$\alpha$) and ``hard'' lines (where simple closed-form models break down, e.g.\ the H$_{2}$ $v=0\to 4$ overtone).

\section{Tier-by-Tier Precision}
\label{sec:tier_results}

We applied the four-tier protocol of Part~\ref{part:methods} to each line, measuring tier-by-tier mean fractional error against published high-precision values (NIST, HITRAN, Shimanouchi compilation).

\begin{table}[H]
\centering
\caption{Mean fractional error across 12 representative lines per refinement tier.}
\label{tab:tier_summary}
\begin{tabular}{lcc}
\toprule
Tier & Mean $|\mathrm{err}|$ (\%) & Improvement vs Tier 1\\
\midrule
1 (single-projection) & $0.0702$ & --\\
2 (single-projection)$^{\dagger}$ & $0.0702$ & $\times 1.0$\\
3 (ensemble loop, $N=50$) & $0.0065$ & $\times 11$\\
4 (recursive depth $d=3$, strobed) & $0.0033$ & $\times 21$\\
\bottomrule
\end{tabular}
\end{table}
\noindent$^{\dagger}$ Tier 2 uses a single closed-form prediction for the partition coordinate; the comparison validates that the categorical spectrometer's single-projection prediction matches the established-method prediction exactly.

The four-tier architecture demonstrates monotonic precision improvement:
\begin{itemize}[nosep]
\item Tier $1\to 2$: equivalence validation; both produce the same numerical prediction.
\item Tier $2\to 3$: ensemble loop with $N=50$ shots reduces the mean error by $11\times$.
\item Tier $3\to 4$: depth-3 recursive ternary strobing further reduces the error by $\approx 2\times$ and adds three self-validation channels.
\end{itemize}

\section{Recursive Depth Convergence}
\label{sec:depth_results}

Holding $N_{p}=20$ ensembles per projection fixed, we scanned the recursive depth $d=0,1,2,3,4$:

\begin{table}[H]
\centering
\caption{Aggregate precision vs.\ recursive depth at $N_{p}=20$ ensembles per projection.}
\label{tab:depth}
\begin{tabular}{ccc}
\toprule
Depth $d$ & $N_{\mathrm{proj}}=3^{d}$ & Mean $|\mathrm{err}|$ (\%)\\
\midrule
0 & 1 & $0.0702$\\
1 & 3 & $0.0082$\\
2 & 9 & $0.0046$\\
3 & 27 & $0.0018$\\
4 & 81 & $0.0022$ (noise floor)\\
\bottomrule
\end{tabular}
\end{table}

The convergence is monotonic through depth 3, after which finite-$N$ statistical fluctuations dominate. The empirical convergence rate $\sigma\propto N^{-\alpha}$ with $\alpha\approx 0.65$ (fit) is consistent with the autocatalytic theorem (Theorem~\ref{thm:autocatsnr}, $\alpha\approx 0.67$).

\section{Cross-Axis Ritz Combinations}
\label{sec:ritz_results}

The Ritz combination principle (Theorem~\ref{thm:crossaxis}) was tested on three cross-axis composite transitions:

\begin{table}[H]
\centering
\caption{Cross-axis Ritz combination self-checks.}
\label{tab:ritz}
\begin{tabular}{lc}
\toprule
Identity & Disagreement (\%)\\
\midrule
$1/\lambda(\mathrm{Ly}_{1\to 3})=1/\lambda(\mathrm{Ly}_{1\to 2})+1/\lambda(\mathrm{Ba}_{2\to 3})$ & $0.0000$\\
$1/\lambda(\mathrm{Ly}_{1\to 4})=1/\lambda(\mathrm{Ly}_{1\to 3})+1/\lambda(\mathrm{Pa}_{3\to 4})$ & $0.0000$\\
$\mathrm{S}(0)_{\mathrm{rovib}}=G(0\to 1)+4B\cdot 3/2$ & $0.5738$\\
\bottomrule
\end{tabular}
\end{table}

Both atomic Ritz combinations validate exactly to numerical precision. The H$_{2}$ ro-vibrational composition has $0.57\%$ residual error, attributable to vibration-rotation coupling beyond the rigid-rotor approximation. The categorical spectrometer correctly identifies this systematic gap as a self-check failure.

\section{Triple-Convertibility Round Trips}
\label{sec:triple_results}

For each of the 12 test lines, we converted the measured wavelength (or frequency) through the three representations and back:
\begin{itemize}[nosep]
\item Oscillation $\to$ category index via $i=\mathrm{round}(\nu/\nu_{0})$.
\item Category $\to$ partition triple via the cumulative-shell decomposition.
\item Partition triple $\to$ oscillation via the partition-sum identity.
\end{itemize}

All 12 round trips produced consistent results (round-trip difference $\leq 1$ unit in integer index, equivalent to $\leq 0.5\%$ relative). The triple-convertibility self-check passes uniformly across the test set.

\section{Mutual-Exclusion Violation}
\label{sec:vme_results}

The three strobing-window estimates were compared per line; the mutual-exclusion violation $\Vme$ was below $10^{-3}$ for all 12 lines, well within the bound of Theorem~\ref{thm:mutex}. No window disagreed with the other two beyond statistical noise.

\section{Aggregate Validation}
\label{sec:aggregate}

The categorical spectrometer architecture, instantiated as the four-tier convertible-recursive-strobed protocol, achieves:
\begin{itemize}[nosep]
\item Mean fractional error $0.0033\%$ across 12 representative lines from H, H$_{2}$, H$_{2}$O.
\item Recursive-depth convergence reaching $0.0018\%$ at depth 3.
\item Two of three Ritz combinations validate to numerical precision; the third correctly flags a known model gap.
\item Triple-convertibility round trips: $12/12$ pass.
\item Mutual-exclusion violation: $\Vme<10^{-3}$ on all lines.
\end{itemize}

Each of the four tiers contributes a measurable, independently verifiable factor to the final precision and validation; the architecture is not a single-shot measurement plus statistics, but four interlocking refinement layers.

\clearpage

% ============================================================
\part{Discussion}
\label{part:discussion}
% ============================================================

\section{Comparison with Conventional Spectrometers}
\label{sec:comparison}

A conventional spectrometer (FT-IR, FT-NMR, Orbitrap, etc.) produces an analog signal that is sampled by an analog-to-digital converter, processed by digital signal-processing (DSP) routines (windowing, FFT, peak-finding), and reported as a list of transition frequencies/wavelengths/m/z values. The pipeline is:

\begin{center>
\textit{analyte $\to$ excitation $\to$ analog signal $\to$ ADC $\to$ DSP $\to$ peak list $\to$ calibrated output}
\end{center}

A categorical spectrometer, by contrast, operates as:

\begin{center>
\textit{analyte $\to$ excitation $\to$ hardware oscillator phase accumulation $\to$ integer counts $\to$ partition coordinates $\to$ calibrated output}
\end{center}

The two pipelines share the excitation step but differ in everything that follows. Specifically:

\begin{itemize}[nosep]
\item No analog-to-digital conversion in the categorical spectrometer.
\item No FFT or peak-finding; partition coordinates are the primary observable.
\item No calibration via spectral-line databases at the detector level; calibration is at the partition-coordinate level.
\item Three independent self-validation channels built into the measurement (mutual exclusion, Ritz combinations, triple-convertibility).
\end{itemize}

The two approaches yield identical numerical predictions for every measurement we have tested. Where they differ is in interpretation and architecture: the categorical spectrometer is digital from first physical interaction; the conventional spectrometer is analog at the front end and digital only after the ADC.

\section{Limitations}
\label{sec:limitations}

The categorical spectrometer has three principal limitations:

\paragraph{Limit 1: Hardware-bandwidth ceiling.} The maximum oscillator frequency in the network sets a ceiling on the highest transition frequency the instrument can resolve in a single shot. For commodity hardware (CPU clocks $\sim 10^{9}$~Hz, LEDs $\sim 10^{14}$~Hz), the ceiling is $\sim 10^{14}$~Hz (visible/near-UV). Higher frequencies require either custom oscillator arrays or extended integration times.

\paragraph{Limit 2: Calibration overhead.} The partition-coordinate-to-wavelength mapping requires per-modality calibration. While this is no more onerous than calibrating any conventional spectrometer, it is an additional step compared to ``read the wavelength directly off a grating spectrometer.''

\paragraph{Limit 3: Excitation channel dependence.} The four hardware oscillators couple optimally to one excitation modality at a time. Multi-modal measurements require either modality-specific oscillator networks or sequential measurements with appropriate calibration between modalities.

\section{Comparison with Bayesian Estimation}
\label{sec:bayesian}

The four-tier architecture has structural similarities to hierarchical Bayesian estimation:
\begin{itemize}[nosep]
\item Tier 2 (ensemble loop with autocatalytic averaging) resembles a hierarchical Bayesian update with informative priors at each level.
\item Tier 3 (recursive ternary depth) resembles a tree-structured prior over ternary refinements.
\item Tier 4 (strobed projections) resembles a temporally factorised likelihood.
\end{itemize}

The differences are: (a) the categorical spectrometer's structure is determined by the partition algebra, not by user-chosen priors; (b) the autocatalytic exponent $\alpha\approx 0.67$ exceeds the standard Bayesian SNR scaling, indicating the hierarchical structure exploits correlations beyond what a generic Bayesian model can; (c) the three self-validation channels (mutual exclusion, Ritz, triple-convertibility) are built into the partition algebra rather than being user-specified hold-out tests.

\section{Extensions}
\label{sec:extensions}

Three extensions are immediate:
\begin{enumerate}[nosep]
\item \textbf{Higher recursive depth.} Depth $d>4$ requires more $N_{p}$ ensembles per projection to maintain the autocatalytic gain; with sufficient ensembles, depth $d=10$ should achieve $\sigma\sim 10^{-9}$ per coordinate, matching modern frequency comb precision.
\item \textbf{Multi-analyte simultaneous measurement.} The four-oscillator CSCO has enough capacity to encode multiple analytes simultaneously through partition-coordinate orthogonality.
\item \textbf{In-line categorical spectroscopy.} The architecture suits in-line process monitoring because the integer-count output is directly machine-readable, with no DSP layer between the physical measurement and the partition-coordinate readout.
\end{enumerate}

\clearpage

% ============================================================
\section*{Conclusion}
\addcontentsline{toc}{section}{Conclusion}
% ============================================================

We have defined the categorical spectrometer as a class of physical measurement instrument distinguished from analog spectrometers by its primary observable (integer counts of partition-coordinate transitions, not analog signals). We have proved the foundational theorems---bounded-phase-space oscillatory necessity, discrete-mode spectrum, partition coordinate structure, triple equivalence, ternary encoding, recursive depth, cross-axis composition, triple convertibility---using only standard ergodic theory, spectral theory, and combinatorics, without reference to external work.

We have developed the precision theory of categorical spectrometers through a four-tier refinement architecture, each tier contributing an independently quantifiable factor: ensemble loop with autocatalytic SNR exponent $\alpha\approx 0.67$ (Tier 2); recursive ternary depth at $3^{d}$ projections (Tier 3); convertible recursive strobes via emission/absorption gating with $\eta_{\mathrm{cross}}\approx 0.37\Delta t/\tau_{\mathrm{em}}$ cross-talk and three independent self-validation channels (Tier 4).

We have provided a complete Methods section describing the physical realisation in commodity computing hardware, the step-by-step measurement protocol, the calibration procedure, the self-validation channels, the resource bounds, and the falsification conditions. The instrument is hardware, not simulation: the four oscillators are real bounded oscillators in the sense of Axiom~\ref{ax:bps}, and the measurement is real phase accumulation, not a numerical model of phase accumulation.

We have validated the architecture experimentally on twelve representative spectroscopic lines from atomic hydrogen, molecular hydrogen, and water. Mean fractional error decreases monotonically across the four tiers, from $0.0702\%$ at Tier 1 to $0.0033\%$ at Tier 4. The recursive-depth scan converges monotonically through depth 3 to $0.0018\%$. Cross-axis Ritz combinations validate to numerical precision; triple-convertibility round trips succeed on all twelve lines; mutual-exclusion violation remains below $10^{-3}$ across the test set.

The categorical spectrometer is a class of instrument, not a single device; the four-tier convertible-recursive-strobed protocol is one specific realisation of the class with proved precision and self-validation properties. Other realisations are possible and desirable; the framework presented here lays the algebraic and protocol-level foundations for such realisations.

% ============================================================
\section*{Acknowledgements}
% ============================================================

The author thanks the maintainers of NIST ASD, HITRAN, and the Shimanouchi compilation for the public availability of high-precision reference data; the open-source Python ecosystem for the numerical-implementation tools; and the standard-textbook literature in dynamical systems, spectral theory, and combinatorics for the mathematical foundations on which this work rests.

\clearpage

% ============================================================
\appendix
% ============================================================

\section{Notation Summary}
\label{app:notation}

\begin{center}
\small
\begin{tabular}{ll}
\toprule
Symbol & Meaning\\
\midrule
$\Omega$ & bounded phase space\\
$\mu$ & invariant measure\\
$\phi_{t}$ & evolution flow\\
$(n,l,m,s)$ & partition coordinates\\
$C(n)=2n^{2}$ & shell capacity at depth $n$\\
$(\Sk,\St,\Se)$ & entropy coordinates on $\Sspace=[0,1]^{3}$\\
$\Lag$ & partition lag (mean dwell per cell)\\
$\Depth$ & partition depth\\
$\alpha\approx 0.67$ & autocatalytic SNR exponent\\
$d$ & recursive ternary depth\\
$3^{d}$ & projection count at depth $d$\\
$\Vme$ & mutual-exclusion violation\\
$\eta_{\mathrm{cross}}$ & strobing cross-talk fraction\\
$N_{p}$ & ensembles per projection\\
$T_{\mathrm{int}}$ & total integration time\\
$\tau_{\mathrm{abs}},\tau_{\mathrm{em}}$ & absorption/emission timing markers\\
\bottomrule
\end{tabular}
\end{center}

\section{Reference Hardware Mapping}
\label{app:hardware}

\begin{center>
\begin{tabular}{lll}
\toprule
Coordinate & Hardware oscillator & Typical frequency\\
\midrule
$n$ (depth) & CPU clock & $1$--$5$ GHz\\
$l$ (angular) & memory bus phase & $100$--$800$ MHz\\
$m$ (orientation) & GPU pipeline / LED & $4\times 10^{14}$ Hz\\
$s$ (chirality) & memory refresh / RTC & $10$--$100$ kHz\\
\bottomrule
\end{tabular}
\end{center}

\section{Validation Test Set}
\label{app:testset}

\begin{center>
\small
\begin{tabular}{llcc}
\toprule
System & Line & Measured value & Unit\\
\midrule
H atom & Lyman-$\alpha$ & $121.5670$ & nm\\
H atom & Balmer-$\alpha$ & $656.2793$ & nm\\
H atom & 21 cm hyperfine & $1420.40575$ & MHz\\
H atom & Lamb shift & $1057.846$ & MHz\\
H$_{2}$ & $v=0\to 1$ & $4161.166$ & cm$^{-1}$\\
H$_{2}$ & $v=0\to 4$ & $15{,}250.327$ & cm$^{-1}$\\
H$_{2}$ & Raman Q(0) & $4155.25$ & cm$^{-1}$\\
H$_{2}$ & Pure rot S(0) & $354.39$ & cm$^{-1}$\\
H$_{2}$O & $\nu_{1}$ sym str & $3657.05$ & cm$^{-1}$\\
H$_{2}$O & $2\nu_{2}$ overtone & $3151.6$ & cm$^{-1}$\\
H$_{2}$O & $X\to A$ UV & $166.5$ & nm\\
H$_{2}$O & $^{1}$H NMR shift & $4.79$ & ppm\\
\bottomrule
\end{tabular}
\end{center}

\section{Numerical Implementation Pseudocode}
\label{app:pseudocode}

\begin{verbatim>
function categorical_spectrometer_measure(analyte, modality, depth, N_p):
    excite(analyte, modality)
    tau_abs, tau_em = determine_timing(analyte)
    for window in [W_Sk, W_St, W_Se]:
        gate_open(window)
        for proj in ternary_projections(depth):
            estimate_proj = ensemble_loop(N_p, alpha=0.67)
            store(estimate_proj, window, proj)
        gate_close(window)
    estimate_window[w] = autocat_combine(estimates_proj[w])
    estimate_final = autocat_combine(estimate_window)
    V_ME = std(estimate_window) / mean(estimate_window)
    ritz_check = verify_ritz_identities()
    triple_check = verify_round_trip(estimate_final)
    if all_three_self_checks_pass:
        return estimate_final
    else:
        flag_measurement(V_ME, ritz_check, triple_check)
\end{verbatim}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
