% =====================================================================
%  The Irreducible Boundary of a Bounded Resolvable Space
%  A self-contained theory of individuation, boundary thickness,
%  and the resolution floor.
% =====================================================================
\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[a4paper,margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{bm}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage{microtype}
\usepackage[round,sort&compress]{natbib}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc,decorations.pathmorphing,fit}
\usepackage[colorlinks=true,linkcolor=blue!45!black,
            citecolor=blue!45!black,urlcolor=blue!45!black]{hyperref}
\usepackage{cleveref}

% ---- Theorem environments ----
\newtheoremstyle{thmstyle}{\topsep}{\topsep}{\itshape}{}{\bfseries}{.}{.5em}{}
\theoremstyle{thmstyle}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{axiom}{Axiom}
\newtheorem{principle}[theorem]{Principle}
\newtheorem{example}[theorem]{Example}
\newtheorem{construction}[theorem]{Construction}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{notation}[theorem]{Notation}

\crefname{axiom}{Axiom}{Axioms}
\Crefname{axiom}{Axiom}{Axioms}
\crefname{principle}{Principle}{Principles}
\Crefname{principle}{Principle}{Principles}

% ---- Shorthand ----
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Space}{\Omega}
\newcommand{\Bdy}{\partial}
\newcommand{\interior}{\mathrm{int}}
\newcommand{\cl}{\mathrm{cl}}
\newcommand{\diam}{\mathrm{diam}}
\newcommand{\dist}{\mathrm{dist}}
\newcommand{\Leb}{\mathcal{L}}
\newcommand{\Haus}{\mathcal{H}}
\newcommand{\thick}{\beta}
\newcommand{\reso}{\delta}
\newcommand{\floor}{\thick}
\newcommand{\Part}{\mathscr{P}}
\newcommand{\cell}{\mathsf{c}}
\newcommand{\eps}{\varepsilon}
\newcommand{\Neg}{\mathcal{N}}
\newcommand{\Ver}{\mathsf{V}}
\newcommand{\norm}[1]{\lVert#1\rVert}
\newcommand{\abs}[1]{\lvert#1\rvert}
\newcommand{\dd}{\,\mathrm{d}}

\title{\textbf{The Irreducible Boundary of a Bounded Resolvable Space:\\[2pt]
Individuation, Boundary Thickness, and the Resolution Floor}}

\author{%
Kundai Farai Sachikonye\\
\small Technical University of Munich\\
\small Chair of Proteomics and Bioanalytics\\
\small \texttt{kundai.sachikonye@wzw.tum.de}}

\date{\today}

\begin{document}
\maketitle

% =====================================================================
\begin{abstract}
\noindent
We study \emph{individuation} in a bounded resolvable space: the act of
separating a part from the whole when no element of the space is
distinguished in advance. We prove that, in a bounded space carrying a
finite resolution, individuation can proceed only by negation---a part is
fixed as the complement of everything it is not---and that every such
separation deposits content of strictly positive measure into a boundary.
This content, the \emph{resolution floor} $\thick>0$, is the central
invariant of the paper. We establish a \emph{Boundary--Thickness Theorem}
(no separator between a part and its complement has zero content), a
\emph{Non-Instantaneity Theorem} (a separation that yields a genuine part
cannot be completed in zero steps, and what the process leaves untouched is
an irreducible residue), and a \emph{Detectability Theorem} (a part is
distinguishable exactly when its own scale exceeds its boundary thickness;
sub-floor parts are engulfed by their separator). The classical sorites is a
corollary: there is no sharp count at which an aggregate becomes
distinguishable, because the separating boundary has positive width and
distinguishability is emergence from that width. We give three independent
derivations of the floor---representational, cardinality-theoretic, and a
cost argument by which compressing a separator to zero content diverges---and
prove they bound the same constant. We then identify the deepest obstruction:
to \emph{verify} that a part is what it is requires deciding its membership
in the very collection that constitutes it (its negation), a self-referential
question with no consistent value. The irreducible residue is therefore not a
locatable set but a self-referential impossibility, proved here by a
self-contained diagonal argument; it cannot be exhibited, only bounded below
by $\thick$. From this we obtain a \emph{Non-Return Theorem}---the count of
committed separations is monotone non-decreasing under every traversal, so no
process returns to a prior individuation state, and resemblance never entails
identity---and an \emph{Inquiry--Cessation Theorem}: thinning a separator
toward zero content has diverging cost and bounded return, so a finite agent
halts at the floor, and knowledge thereafter attaches to bounded wholes and
never to their sub-floor parts. Every claim is proved from stated definitions;
uncertain material has been omitted.
\end{abstract}

\noindent\textbf{Keywords:} bounded resolvable space; individuation;
negation; topological boundary; boundary thickness; resolution floor; sorites
paradox; vagueness; self-reference; diagonal argument; irreducible residue;
monotone individuation.

\tableofcontents

% =====================================================================
\section{Introduction}
\label{sec:intro}
% =====================================================================

\subsection{The problem of individuation without a selector}

A bounded space contains no part until one is made. To speak of a part $A$ is
to have separated $A$ from its complement; but if no element of the space is
distinguished in advance---if there is no privileged point, no selector, no
prior ``main actor''---then the first question is how a part comes to be
individuated at all. This paper takes that question as primary and answers it
with a single device, negation: a part is fixed not by being chosen but by
being marked off from \emph{everything it is not}. The complement requires no
selector, because it is the whole minus the part and is given as soon as the
part is; and the part is recovered as the one thing the complement omits. We
prove that this is essentially the \emph{only} selector-free individuation
(\Cref{sec:negation}), and we trace its consequences.

\subsection{Thesis}

Under one hypothesis---that the space is bounded and finitely resolvable
(\Cref{ax:brs})---we prove the following chain.

\begin{enumerate}[label=(\roman*),leftmargin=2.2em]
\item \emph{Individuation is by negation.} A part is the complement of its
negation; this is the unique selector-free way to obtain a determinate part
from an undifferentiated whole.
\item \emph{Every separation has positive thickness.} No part can be cut from
its complement by a separator of zero content. The separator carries content
$\ge\thick>0$, the resolution floor (\Cref{sec:thickness}).
\item \emph{Separation is non-instantaneous, and leaves a residue.} A
separation completed in zero steps yields no part; an actual separation
occupies the process and leaves an untouched remainder
(\Cref{sec:negation}).
\item \emph{Detection is emergence from the boundary.} A part is
distinguishable exactly when its scale exceeds its own boundary thickness;
below that scale it is engulfed (\Cref{sec:detect}). The sorites paradox is a
corollary.
\item \emph{Verification is self-referentially obstructed.} To verify a part
is to decide its membership in the collection that constitutes it; this has no
consistent value, so the residue is irreducible---bounded below by $\thick$
yet not exhibitable as a set (\Cref{sec:residue}).
\item \emph{Individuation does not return.} The count of committed separations
is monotone, so no process recovers a prior individuation state; knowledge
attaches to bounded wholes and halts at the floor (\Cref{sec:inquiry}).
\end{enumerate}

\subsection{On the word ``knowledge''}

We use \emph{knowledge} in a single, fully operational sense throughout: to
know a part is to be able to verify it, and to verify a part is to decide,
against its complement, that it is the part it is. No other notion of
knowledge is invoked, and nothing in the paper depends on any informal
reading of the word. With this definition, ``the boundary of knowledge'' is
literally the boundary of the verifying separation, and the theorems about
boundaries are theorems about knowledge in this exact sense.

\subsection{Why ``thickness''}

The everyday and the formal meet at the sorites. An aggregate built one unit
at a time is, at first, not distinguishable as an aggregate; at some size it
is; yet no single unit is the one that makes it so. Classically this is a
paradox. We resolve it by inverting the reading: a sub-threshold collection is
indistinguishable not because it is small in the absolute but because its
\emph{separating boundary is thick relative to it}---the layered separator one
must resolve to mark it off from everything else is wider than the collection,
so it is engulfed by its own boundary and never stands out. It becomes
distinguishable when it is large compared with that separator. There is no
critical unit because distinguishability is emergence across a band of
positive width, not the crossing of a point; the width of the band is the
floor $\thick$ (\Cref{thm:detectability}, \Cref{cor:sorites}).

\subsection{Why ``bounded''}

Boundedness is the source of the floor, not a convenience. On an unbounded
space a part may be separated from its complement by an arbitrarily thin
frontier, and structure may be resolved to arbitrary depth at fixed cost. It
is the finiteness of the space, together with the finiteness of the resolution
any embedded agent can realise, that forbids the zero-thickness cut. We make
the hypothesis exactly strong enough to force the floor and no stronger, and
we mark, at each step, where boundedness is used (\Cref{prop:bdd-necessary}).

\subsection{Relation to existing work}

The paper is pure mathematics: topology and measure theory~\citep{kelley1955,%
munkres2000,halmos1950,federer1969,mattila1995,engelking1995,hurewicz1941},
order and fixed-point theory~\citep{birkhoff1967,daveypriestley2002,%
tarski1955}, and self-reference~\citep{russell1903,cantor1891,tarski1936,%
godel1931,lawvere1969}. The sorites and vagueness have substantial
literatures---degree and fuzzy theories~\citep{zadeh1965}, supervaluation
\citep{fine1975,keefe2000}, epistemicism~\citep{williamson1994}, and the
classical analysis of~\citet{black1937}; see~\citet{hyderaffman2018} for a
survey. Our treatment is none of these: it is not a logic of partial truth,
nor a claim that a sharp boundary exists but is unknown, but a
\emph{geometric} statement that the separator is a set of positive content
whose verification is self-referentially obstructed. The self-reference
results we use are proved here in self-contained form; the classical results
of~\citet{russell1903,cantor1891,tarski1936} are cited only as the shape the
argument shares. A reasonable reader will recognise that the abstract
hypotheses below are satisfied by a range of concrete systems; we make no such
identification and prove everything of the abstract space alone.

\subsection{Method}

Every substantive statement is a definition, a theorem proved here, or a
citation of a standard mathematical result compared against such a theorem.
Where a tempting claim could not be proved at this standard, it has been
removed rather than weakened. The development is uniform: a single hypothesis,
a single invariant $\thick$, and a chain of theorems that return to it from
five directions---topological, representational, cardinality-theoretic,
self-referential, and order-theoretic---each shown to bound the same constant.

\subsection{Organisation}

\Cref{sec:setting} fixes the bounded resolvable space and its hypothesis.
\Cref{sec:negation} proves that individuation must be by negation and that it
is non-instantaneous, yielding the residue. \Cref{sec:thickness} proves the
Boundary--Thickness Theorem. \Cref{sec:floor} gives the three agreeing
derivations of the floor. \Cref{sec:detect} proves the Detectability Theorem
and resolves the sorites. \Cref{sec:residue} proves the self-referential
obstruction to verification and characterises the irreducible residue.
\Cref{sec:inquiry} proves the Non-Return and Inquiry--Cessation Theorems.
\Cref{sec:scope} states scope and concludes.
