
% =====================================================================
\section{Non-self-verifiability and the irreducible residue}
\label{sec:residue}
% =====================================================================

The previous sections show the separator has positive content and that this
content silences sub-floor regions. We now reach the deepest obstruction: the
separator cannot be \emph{verified} from within. To verify a part is to decide it
against the totality that constitutes it---its negation-sequence---and that
decision is self-referential, with no consistent value. We prove this by a
self-contained diagonal argument (\Cref{thm:diagonal}), deduce that the
irreducible residue is not a locatable set but a self-referential impossibility
(\Cref{thm:residue-not-measure}), and conclude that the floor can only ever be
bounded below, never exhibited (\Cref{cor:floor-only-bounded}).

\subsection{Verification as deciding membership in the negation}

Recall (\Cref{def:negation}) that $A$ is constituted as $\Space\setminus\Neg(A)$:
its identity is fixed by the totality $\Neg(A)=\Space\setminus A$ of what it is
not. We make verification precise as the decision that constitutes this identity.

\begin{definition}[Verifier]
\label{def:verifier}
A \emph{verifier} for the parts of $\Space$ is a map $\Ver$ that, given a part
$X$ and a candidate constituting totality $T$, returns $\Ver(X,T)\in\{0,1\}$,
intended to read $1$ iff $X$ is exactly the part constituted by withholding $T$,
i.e.\ iff $X=\Space\setminus T$. To \emph{verify $A$} is to evaluate
$\Ver(A,\Neg(A))$: to decide whether $A$ is the part its own negation-sequence
constitutes.
\end{definition}

\begin{remark}[Why verification is membership in the negation]
\label{rem:ver-membership}
By \Cref{def:negation}, $A=\Space\setminus\Neg(A)$ exactly when $A$ and $\Neg(A)$
partition $\Space$; deciding this is deciding, of the constituting totality
$\Neg(A)$, whether what it omits is precisely $A$---a question about $A$ posed
entirely in terms of the collection $\Neg(A)$ that is defined as
``everything but $A$.'' Verification of $A$ is thus inseparable from a question
of self-membership: $A$ figures, through its own definition, inside the totality
against which it is checked.
\end{remark}

\subsection{A self-contained diagonal obstruction}

We now prove, without invoking any external impossibility theorem, that no
verifier decides every part correctly. The argument is a diagonal one, in the
shape of the classical results of~\citet{russell1903,cantor1891,tarski1936} and
the general form of~\citet{lawvere1969}, but it is self-contained: it uses only
the definitions above.

Two ingredients are needed before the diagonal can be run, and we isolate them as
an explicit hypothesis, because whether they hold is exactly the delicate point.
The diagonal collection
\[
D\;:=\;\bigl\{\,X : \Ver(X,\Neg(X))=0\,\bigr\}
\]
is, a priori, only a \emph{collection of parts indexed by the verifier's
behaviour}; for the verifier to be asked about $D$ in the sense of the theory, $D$
must itself be an admissible object of the same kind---a region of $\Space$ on
which $\Ver$ is defined.

\begin{definition}[Self-applicable verifier]
\label{def:self-applicable}
A verifier $\Ver$ is \emph{self-applicable} if
\begin{enumerate}[label=\textup{(SA\arabic*)},leftmargin=3em]
\item\label{sa:total} it is total on the regions of $\Space$ (it returns a value
in $\{0,1\}$ on every region and its negation), and
\item\label{sa:closed} the diagonal collection $D$ above is itself a region of
$\Space$---a Borel set on which $\Ver(\,\cdot\,,\Neg(\,\cdot\,))$ is defined---so
that $\Ver(D,\Neg(D))$ is a legal query.
\end{enumerate}
\end{definition}

\begin{theorem}[Diagonal obstruction to verification]
\label{thm:diagonal}
No self-applicable verifier exists. Equivalently, there is no verifier $\Ver$
that is total on regions \ref{sa:total} and closed under its own diagonal
\ref{sa:closed}: for any such $\Ver$ the value $\Ver(D,\Neg(D))$ cannot be
consistently assigned, since $\Ver(D,\Neg(D))=1$ and $\Ver(D,\Neg(D))=0$ each
entail the other. In particular no verifier satisfying \ref{sa:closed} returns
$1$ on every region.
\end{theorem}

\begin{proof}
Suppose, for contradiction, a self-applicable verifier $\Ver$ exists. By
\ref{sa:closed} the collection $D$ is a region and $\Ver(D,\Neg(D))$ is a legal
query; by \ref{sa:total} it has a value. $D$ collects exactly those parts $X$ for
which the verifier returns $0$ on the self-referential query $\Ver(X,\Neg(X))$.
Now ask the verifier about $D$ itself.

If $\Ver(D,\Neg(D))=1$, then $D$ is declared constituted by its own negation;
but by the defining property of $D$, membership-defining condition for $D$ is
$\Ver(X,\Neg(X))=0$, so $D$ satisfying its own membership condition would require
$\Ver(D,\Neg(D))=0$. Reading $D$ through its own definition therefore forces
$\Ver(D,\Neg(D))=0$, contradicting the assumed value $1$.

If instead $\Ver(D,\Neg(D))=0$, then $D$ meets its own membership condition, so
$D$ is among the parts collected by $D$, i.e.\ $D$ is constituted as the part its
negation withholds, which is the assertion $\Ver(D,\Neg(D))=1$, contradicting the
assumed value $0$.

Both values are self-refuting, so no consistent value exists for
$\Ver(D,\Neg(D))$. This contradicts \ref{sa:total}, which requires a value. Hence
no verifier satisfies both \ref{sa:total} and \ref{sa:closed}; in particular none
satisfying \ref{sa:closed} returns $1$ on every region. This is the diagonal
obstruction.
\end{proof}

\begin{remark}[On the closure hypothesis \ref{sa:closed}]
\label{rem:self-app}
The theorem is exactly as strong as \ref{sa:closed}, and we are careful not to
claim more than we prove. \ref{sa:closed} asserts that the behaviour-defined
collection $D$ is again a region of $\Space$ on which the verifier is defined.
This is the genuine content of ``self-application'': the verifier can be asked
about an object built from its own verdicts. Two remarks on its status.

\emph{Why embedding makes it natural.} An agent that verifies regions of $\Space$
while being itself a part of $\Space$ is a plausible setting in which a
description of the agent's own verifying behaviour is realised as a region---the
agent's verdicts leave a trace in the same space they range over, and a region
selected by those verdicts is a region of that space. In such a setting
\ref{sa:closed} holds and the obstruction bites.

\emph{Why it is not automatic.} Nothing in \Cref{ax:brs} forces a
behaviour-defined collection to be Borel-encodable as a region. The classical
diagonal results show the shape of the argument but not that its preconditions are
met here: Russell's construction lives in unrestricted comprehension, which the
measurable setting does not grant; and the general fixed-point
theorem of~\citet{lawvere1969} delivers the contradiction only given a
point-surjection in a cartesian-closed category---structure that an embedding
\emph{may} supply but that must be exhibited, not assumed. We therefore treat
\ref{sa:closed} as a hypothesis, true in any model where the agent's verifying
behaviour is itself realisable as a region, and we mark every downstream claim
that depends on it. Discharging \ref{sa:closed} in full generality---constructing
the measurable encoding of parts on which $\Ver$ acts as a point-surjection---is a
separate undertaking we do not claim to have carried out.
\end{remark}

\subsection{The residue is not a set under the closure hypothesis}

The diagonal obstruction lets us characterise the irreducible residue, under the
same closure hypothesis \ref{sa:closed} that the obstruction requires. One might
hope that the un-verified remainder of an individuation is a definite set---small,
locatable, perhaps measurable---which a finer effort could in principle exhibit
and remove. We show that this hope is incompatible with \ref{sa:closed}: a
locatable residue would let one assemble a total verifier closed under its own
diagonal, which \Cref{thm:diagonal} forbids.

\begin{definition}[Irreducible residue]
\label{def:irreducible}
The \emph{irreducible residue} of the individuation of $A$ is the obstruction to
deciding $\Ver(A,\Neg(A))$: the content that an individuation cannot certify
about itself. It is \emph{exhibitable} if there is a realisable region $E$ with
$\mu(E)>0$ such that verifying \emph{every} region reduces to verifying it on
$\Space\setminus E$ together with a separate, always-terminating certification of
the residue localised in $E$.
\end{definition}

\begin{theorem}[The irreducible residue is not exhibitable, under \ref{sa:closed}]
\label{thm:residue-not-measure}
Let \Cref{ax:brs} hold and suppose the closure hypothesis \ref{sa:closed}---that
a behaviour-defined collection of regions is again a region. Then the irreducible
residue is not exhibitable: there is no realisable region $E$ of positive measure
whose separate certification completes the verification of every region. Under
\ref{sa:closed} the residue is a self-referential impossibility, not a locatable
set.
\end{theorem}

\begin{proof}
Suppose the residue were exhibitable by some realisable $E$ with $\mu(E)>0$:
certifying the residue localised in $E$ always terminates and, combined with
verification on $\Space\setminus E$, completes the verification of every region.
This combination is a verifier $\Ver^\ast$ that is total on regions
\ref{sa:total}. By the standing hypothesis \ref{sa:closed}, the diagonal
collection $D$ of $\Ver^\ast$ is again a region; so $\Ver^\ast$ satisfies both
\ref{sa:total} and \ref{sa:closed}, i.e.\ is self-applicable. \Cref{thm:diagonal}
shows no such verifier exists---a contradiction. Hence no exhibiting $E$ exists:
under \ref{sa:closed} the residue cannot be localised to a positive-measure set
and certified away. It is the standing impossibility of the self-referential
decision, not a region awaiting finer treatment.
\end{proof}

\begin{remark}[What survives if \ref{sa:closed} fails]
\label{rem:if-sa-fails}
If the closure hypothesis fails---if the agent's verifying behaviour is not
realisable as a region---then \Cref{thm:diagonal} and \Cref{thm:residue-not-measure}
do not apply, and the present section establishes nothing about exhibitability. The
paper's quantitative core is untouched by this: the floor $\thick>0$ is already
established three independent ways in \Cref{sec:thickness,sec:floor}, before any
diagonal argument is invoked, and the detectability, locating, and non-return
results of \Cref{sec:detect,sec:locating,sec:inquiry} rest on those derivations
alone. What rides on \ref{sa:closed} is solely the \emph{upgrade} from ``the floor
is positive'' to ``the residue is not a locatable set, only boundable below.'' We
state that upgrade as conditional and claim no more.
\end{remark}

\begin{corollary}[The floor can only be bounded below, under \ref{sa:closed}]
\label{cor:floor-only-bounded}
The resolution floor $\thick$ admits a strict positive lower bound
\textup{(}$\thick\ge\mu_{\min}>0$, \Cref{thm:thickness}\textup{)}
unconditionally. If in addition the closure hypothesis \ref{sa:closed} holds,
then $\thick$ has no exact exhibition as $\mu(R)$ for a definite residue set $R$:
every attempt to write the residue as the measure of a located set $R$ supplies
an exhibitable $E=R$, contradicting \Cref{thm:residue-not-measure}. In that case
the floor is known only as the positive quantity below which no individuation can
certify itself.
\end{corollary}

\begin{proof}
The lower bound is \Cref{thm:thickness}. If $\thick=\mu(R)$ for a realisable $R$
of positive measure, then $R$ exhibits the residue, contradicting
\Cref{thm:residue-not-measure}. Hence no such exhibition exists, and $\thick$ is
characterised solely as a positive lower bound on un-certifiable content.
\end{proof}

\begin{remark}[The recognition gap]
\label{rem:recognition-gap}
\Cref{thm:residue-not-measure}, where it applies, has a reading in terms of the
relation between an agent and the part it would verify. To certify $A$ completely, the agent must
decide $A$ against the totality $\Neg(A)$ that constitutes $A$; but the agent,
being embedded, is itself within that totality, so a complete certification would
require it to occupy both sides of the constituting partition at once---to be at
once the part verified and a portion of the totality against which it is
verified. The residue is the irreducible gap that keeps the two apart. Closing
the gap would collapse the distinction on which verification depends; the floor
$\thick$ is the measure of how far the gap can never be closed. This is the
precise sense in which the boundary of knowledge is constitutive, not contingent:
the un-knowable remainder is the very condition under which anything is
individuated at all.
\end{remark}
