% =====================================================================
%  On the Necessary Substructures of Finite Contact Graphs
%  A conditional-existence account of why a finite world is structured.
% =====================================================================
\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[a4paper,margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{bm}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage[round,sort&compress]{natbib}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc,decorations.pathmorphing,fit}
\usepackage[colorlinks=true,linkcolor=blue!45!black,
            citecolor=blue!45!black,urlcolor=blue!45!black]{hyperref}
\usepackage{cleveref}

% ---- Theorem environments ----
\newtheoremstyle{thmstyle}{\topsep}{\topsep}{\itshape}{}{\bfseries}{.}{.5em}{}
\theoremstyle{thmstyle}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{axiom}{Axiom}
\newtheorem{principle}[theorem]{Principle}
\newtheorem{construction}[theorem]{Construction}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{notation}[theorem]{Notation}

\crefname{axiom}{Premise}{Premises}
\Crefname{axiom}{Premise}{Premises}
\crefname{principle}{Principle}{Principles}
\Crefname{principle}{Principle}{Principles}

% ---- Shorthand ----
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Gph}{\Gamma}                 % the contact graph
\newcommand{\V}{V}                        % vertex set
\newcommand{\E}{E}                        % edge set
\newcommand{\wt}{w}                       % edge weight
\newcommand{\floorc}{\beta}               % the floor
\newcommand{\Nbr}{N}                      % neighbourhood
\newcommand{\co}{\complement}             % complement
\newcommand{\cut}{\partial}               % edge boundary / cut
\newcommand{\Walk}{W}                     % a walk
\newcommand{\gate}{g}                     % gating function
\newcommand{\Agent}{A}                    % agent subgraph
\newcommand{\Res}{\rho}                   % residual invariant
\newcommand{\Quot}{Q}                     % quotient graph
\newcommand{\rec}{M}                      % committed record
\newcommand{\Ver}{\mathsf{V}}             % verifier
\newcommand{\med}{\mathsf{M}}             % the medium / whole
\newcommand{\idres}{\rho_{\!*}}            % identification residual
\newcommand{\abs}[1]{\lvert#1\rvert}
\newcommand{\norm}[1]{\lVert#1\rVert}
\newcommand{\restr}{\!\restriction\!}

\title{\textbf{On the Necessary Substructures of Finite Contact Graphs:\\[2pt]
A Conditional-Existence Account of Why a Finite World Is Structured}}

\author{%
Kundai Farai Sachikonye\\
\small Technical University of Munich\\
\small AIMe Registry for Artificial Intelligence\\
\small \texttt{kundai.sachikonye@wzw.tum.de}}

\date{\today}

\begin{document}
\maketitle

% =====================================================================
\begin{abstract}
\noindent
We ask a structural question---\emph{why is a world structured at all?}---and
answer it with a single hypothesis about graphs. A \emph{contact graph} is a
finite weighted graph whose vertices are the parts that have been individuated
within a whole and whose edges carry the cost of separating one part from
another. We first show that the central constant of the theory need not be
assumed: from the two facts that a thing is a proper part of a whole and that the
whole is \emph{non-completable} \textup{(}no finite stage exhausts it\textup{)},
we prove a \emph{Floor from Infinitude} theorem---the comparison identifying a
thing against the rest can never be completed, so every thing carries a strictly
positive irreducible residual. The positive floor is thus derived, not posited;
it is the trace the non-completable whole leaves on each of its finite parts. With
the floor in hand we proceed from a single working premise---that a finite contact
graph with this positive floor exists---and derive, as a chain of existence
theorems, a ladder of substructures any such graph must contain. We prove:
\textbf{(T0)} a strictly positive \emph{floor} on every separating boundary---no
two parts are adjacent at zero cost; \textbf{(T1)} that each vertex is fixed only
by its complement-neighbourhood, so that individuation must proceed by negation
and admits no selector without regress; \textbf{(T2)} that the
\emph{cut-residual}---the inter-part boundary content---is bounded below and is
invariant under every relabelling (isomorphism) of the graph, while no single
vertex realises it, so the residual is a region and never a point; \textbf{(T3)}
that this residual is the conserved quantity under all re-encodings of the
graph's vertices, the unique invariant from which the relabellable surface
derives; \textbf{(T4)} that every connected subgraph carries its own such
invariant under its automorphisms; \textbf{(T5)} that selecting a walk on the
graph requires an edge-gating function, without which the next step is
undetermined; \textbf{(T6)} that the committed-edge count along any walk is
monotone, so revisiting a vertex is never returning to a state and resemblance
never entails identity; \textbf{(T7)} that a coherent family of subgraphs acting
as one admits a single gating function on their quotient graph, and only one;
and \textbf{(T8)} that, because the subgraph invariant is complete, contact
graphs are specifiable and hence instantiable---authored finite structures
exist---subject to three bounds we prove: their walks are not foreseeable from
their specification, an authored copy is a distinct structure, and authorship is
limited to the resolution of the floor. We then give the operational reading: a
\emph{measurement} is a reshuffling of the graph that conserves the medium, so it
tightens a part's boundary while leaving its interior residue---its
identity---untouched; resolving a thing's identity to a point would require
exhausting the non-completable whole, which reshuffling conserves, so identity is
structurally unreachable, and point-identity is refuted in two independent
ways---resolution never saturates and independent lines of measurement never
become redundant. We close with a Master Theorem: every substructure T0--T8 is
forced for finite graphs in a non-completable whole and is \emph{not} forced for
completable or zero-floor graphs, and the whole ladder is equivalent to the
identifiability of any one thing at all. Structure is therefore neither imposed on a world
nor intrinsic to it independent of its parts; it is the necessary form a world
must take to be resolved by finitely many parts within a whole they cannot
exhaust. \emph{The floor is the trace of the whole on its parts; structure is the
shadow of finitude meeting the inexhaustible.} Every claim is proved from the
stated premises; uncertain material has been omitted, and the interpretations a
reader may attach---of parts as agents, residuals as conserved truth, the medium
as the whole, measurement as inquiry---are confined to clearly marked remarks on
which no theorem depends.
\end{abstract}

\noindent\textbf{Keywords:} contact graph; finite weighted graph; edge-weight
floor; individuation by negation; cut invariant; graph isomorphism invariant;
gated walk; monotone walk; quotient graph; conditional existence; necessary
substructure; authored structure.

\tableofcontents

% =====================================================================
\section{Introduction}
\label{sec:intro}
% =====================================================================

\subsection{The question}

Why is a world structured at all? One might expect a world to be either a
featureless continuum, in which nothing is marked off from anything else, or an
arbitrary heap, in which everything is marked off from everything else at no
cost. Neither is what is observed: parts are distinguished, but their
distinction is bounded---there is a cost to telling one part from another, and
the parts compose into a definite architecture. We argue that this architecture
is not contingent. It is forced, in full, by a single fact: that the parts are
\emph{finite in number} and \emph{finitely separated}. The structure of a world
is the shadow its finiteness casts.

We make this precise in the language of graphs, because every relation of
``this part, separated from that part'' is natively an edge between two
vertices, and any architecture of parts is some graph. The object of the paper
is therefore a single graph---a \emph{contact graph}---whose vertices are the
individuated parts of a whole and whose weighted edges record the cost of the
separations between them. We pose one hypothesis about this graph and derive
everything else.

\subsection{Method: conditional existence}

The paper proves no empirical claim. It establishes a conditional: \emph{if}
anything is identifiable at all within a non-completable whole, \emph{then} a
determinate ladder of substructures must exist. We first derive, from
identifiability and non-completability, that the central constant---the positive
floor---is forced rather than assumed (\Cref{sec:ground}); the working premise
below then becomes a consequence, and the remaining rungs are existence theorems
of the form ``under this premise, the following is forced.'' We assume nothing
about what the vertices \emph{are}; we assume only that a thing can be told from
the rest of an inexhaustible whole, and we read off what the resulting graph
cannot lack. The deductions are combinatorial and order-theoretic throughout.
Where a tempting claim could not be proved at this standard it has been removed
rather than weakened.

\subsection{The ground premise and the derived floor}

The deepest premise is not the graph but identifiability itself. \Cref{sec:ground}
takes as given only that there is a whole $\med$ that is \emph{non-completable}
(no finite stage exhausts it) and that a \emph{thing} is a proper part of it. From
these we prove the \emph{Floor from Infinitude} theorem
(\Cref{thm:floor-infinitude}): identifying a thing is comparing it against the
rest of the whole, and were that comparison completable the whole would be
finite; since it is not, the comparison cannot finish and every thing carries a
strictly positive irreducible residual. The positive floor is thereby
\emph{derived}---it is the trace the non-completable whole leaves on each finite
part (\Cref{cor:floor-derived})---and the following Finiteness Premise is retained
only as the working hypothesis for the graph development.

\begin{axiom}[Finiteness Premise]
\label{ax:finite}
A \emph{contact graph} is a finite weighted graph $\Gph=(\V,\E,\wt)$ with
$\abs{\V}<\infty$, in which every edge weight is bounded below by a positive
constant: there is $\floorc>0$ with $\wt(e)\ge\floorc$ for all $e\in\E$. We
assume throughout that a contact graph exists and that it is connected. By
\Cref{cor:floor-derived} the positive floor $\floorc$ is not an independent
posit but the derived residual of \Cref{thm:floor-infinitude}; we keep the
premise in this form for the combinatorial development.
\end{axiom}

The premise is almost nothing: a finite graph whose edges cost something---and
even that cost is earned, not assumed. Its consequences are almost everything we
mean by structure.

\subsection{The ladder of necessary substructures}

Under \Cref{ax:finite} we prove the following chain, each rung from the premise
together with those above it.

\begin{enumerate}[label=\textbf{(T\arabic*)},leftmargin=3em,start=0]
\item \emph{The floor exists} (\Cref{sec:floor}): every separating boundary has
weight $\ge\floorc>0$; no two parts are adjacent at zero cost.
\item \emph{Individuation is by negation} (\Cref{sec:negation}): a vertex is
fixed only as the complement of its non-neighbours; no selector fixes it without
regress.
\item \emph{Truth is a residual region} (\Cref{sec:residual}): the cut-residual
is bounded below and realised by no single vertex, hence a region, never a
point.
\item \emph{The residual is the conserved invariant} (\Cref{sec:residual}):
under every relabelling of vertices the residual is preserved; it is the unique
invariant from which the relabellable surface derives.
\item \emph{Each subgraph has a private invariant} (\Cref{sec:invariant}): every
connected subgraph carries its own invariant under its automorphisms.
\item \emph{Selection requires a gate} (\Cref{sec:gating}): a walk is
undetermined without an edge-gating function that biases the next step.
\item \emph{Histories do not return} (\Cref{sec:monotone}): the committed-edge
count is monotone; revisiting a vertex is not returning to a state.
\item \emph{Plurality forces a single quotient gate} (\Cref{sec:quotient}): a
coherent family of subgraphs acting as one admits exactly one gating function on
their quotient.
\item \emph{Authored structures exist, and are bounded} (\Cref{sec:construct}):
the subgraph invariant is complete, so contact graphs are specifiable and
instantiable; their walks are unforeseeable from the specification, an authored
copy is a distinct structure, and authorship reaches only to the floor.
\end{enumerate}

Two further developments frame the ladder. Beneath it, \Cref{sec:ground} derives
the floor from identifiability in a non-completable whole and names the conserved
reference, the \emph{medium}, against which every part is individuated. Atop it,
\Cref{sec:measurement} gives the operational reading: a measurement is a
reshuffling that conserves the medium, so it tightens a part's boundary while
leaving its interior residue---its identity---untouched; identity-resolution would
require exhausting the whole and is therefore unreachable, and point-identity is
refuted both because resolution never saturates and because independent lines of
measurement never become redundant.

\Cref{sec:master} proves the Master Theorem: T0--T8 are forced for finite
graphs and not forced when finiteness or the positive floor is dropped.

\subsection{On the word ``contact''}

We use \emph{contact} in one operational sense: two parts are \emph{in contact}
when the graph carries an edge between them, and the weight of that edge is the
cost of the separation that individuates each from the other. Nothing in the
paper depends on any informal reading of the word; ``contact'' names an edge and
its weight, and the theorems about contact graphs are theorems about finite
weighted graphs in exactly this sense.

\subsection{Why finiteness is the source}

Finiteness is not a convenience of presentation but the origin of every
structure below. On an infinite graph a vertex may be separated from the rest by
edges of vanishing weight, the residual may be driven to zero, and a walk may be
specified to unbounded depth at fixed cost; none of the substructures is then
forced. It is the finitude of the vertex set, together with the positive floor
on edge weights, that forbids the zero-cost separation and compels the ladder.
We make the premise exactly strong enough to force the structures and no
stronger, and we mark, at each rung, where finiteness is used
(\Cref{sec:master}).

\subsection{Relation to existing work}

The paper is pure mathematics. It uses graph theory~\citep{bondymurty2008,%
diestel2017,west2001,bollobas1998}, algebraic and spectral graph
theory~\citep{godsilroyle2001,chung1997,fiedler1973,cheeger1970}, the theory of
walks and Markov chains on graphs~\citep{lovasz1993,levinperes2017}, order and
fixed-point theory~\citep{daveypriestley2002,birkhoff1967,tarski1955},
enumerative combinatorics and partitions~\citep{stanley2012,andrews1976},
self-reference and diagonal arguments~\citep{russell1903,cantor1891,lawvere1969,%
godel1931}, and information theory~\citep{shannon1948,cover2006}. Where the
development meets the sorites and the logic of vagueness we note the
connection~\citep{black1937,williamson1994} without taking any position in that
literature; our treatment is geometric, not a logic of partial truth. A
reasonable reader will recognise that the abstract premise is satisfied by a
range of concrete systems of finitely many bounded parts; we make no such
identification and prove everything of the abstract graph alone. Such
interpretive readings as we offer are confined to remarks labelled accordingly,
on which no theorem depends.

\subsection{Organisation}

\Cref{sec:ground} derives the floor from identifiability in a non-completable
whole and introduces the medium. \Cref{sec:setting} fixes the contact graph and
its derived objects. \Cref{sec:floor} proves T0. \Cref{sec:negation} proves T1.
\Cref{sec:residual} proves T2 and T3. \Cref{sec:invariant} proves T4.
\Cref{sec:gating} proves T5. \Cref{sec:monotone} proves T6, including the temporal
corollary by which a progressing whole forces non-return. \Cref{sec:quotient}
proves T7. \Cref{sec:construct} proves T8. \Cref{sec:measurement} gives the
operational reading---measurement as reshuffling, the unreachability of identity,
and the two-pronged falsification of point-identity. \Cref{sec:master} proves the
Master Theorem and concludes.
